---
title: '01 Prep - `r unlist(strsplit(getwd(), "/"))[8]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "Diana Acosta [Diana.Acosta@osumc.edu]"
  - name: "Hongjun Fu [Hongjun.Fu@osumc.edu]"
  - name: "Ohio State University Wexner Medical Center"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[8], "01_prep.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Acosta Organoid scRNAseq Project](https://github.com/eturkes/acosta-organoid-scRNAseq).*

In this document we prepare the gene count matrix for downstream analysis.
The data here will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[8]``.

```{r}
#    This file is part of acosta-organoid-scRNAseq.
#    Copyright (C) 2022  Emir Turkes, Diana Acosta, Hongjun Fu, Ohio
#    State University Wexner Medical Center
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

packages <- c(
  "conflicted", "Seurat", "SingleCellExperiment", "scuttle", "DT", "DropletUtils", "parallel",
  "ggplot2", "dplyr", "ggrepel", "viridis"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path("..", "utils.R"))

protocol <- c("human", "droplet", "single-cell", "umis") # See `cluster_pipeline` in `utils.R`.
vars_to_regress <- NULL # See `cluster_pipeline` in `utils.R`.
parallel_override <- 0 # See `parallel_plan` in `utils.R`.

# Metadata to plot after dimensionality reduction and clustering.
# Values in list can include "no_legend and/or "no_label" to exclude those.
# -------------------------------------------------------------------------
metadata_to_plot <- vector("list", 2)
names(metadata_to_plot) <- c("sample", "seurat_clusters")
metadata_to_plot$seurat_clusters <- "no_legend"
# -------------------------------------------------------------------------

analysis_no <- 1
data_name <- unlist(strsplit(getwd(), "/"))[8] # Name of dataset.
data_dir <- file.path("..", "..", "data") # Backed up data.
results_dir <- file.path("..", "..", "results")

# Unique cache directory for each analysis number.
# ------------------------------------------------
cache_dir <- file.path("..", "..", "cache", data_name, paste0("0", analysis_no))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# ------------------------------------------------

knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
```

# Merging

```{r}
rds <- file.path(cache_dir, "seurat_merged.rds")
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {

  file_list <- list.files(file.path(cache_dir, "..", ".."), recursive = TRUE)
  if (length(file_list) == 13) {
    file_list <- file_list[-1]
  }
  files <- vector("list", length(file_list))
  for (i in seq_along(files)) {
    files[[i]] <- readRDS(file.path(cache_dir, "..", "..", file_list[i]))
  }

  genes <- Reduce(
    intersect,
    list(
      rownames(files[[1]]), rownames(files[[2]]), rownames(files[[3]]), rownames(files[[4]]),
      rownames(files[[5]]), rownames(files[[6]]), rownames(files[[7]]), rownames(files[[8]]),
      rownames(files[[9]]), rownames(files[[10]]), rownames(files[[11]]), rownames(files[[12]])
    )
  )
  for (i in seq_along(files)) {
    files[[i]] <- files[[i]][genes, ]
  }

  seurat <- merge(
    files[[1]],
    c(
      files[[2]], files[[3]], files[[4]], files[[5]], files[[6]], files[[7]],
      files[[8]], files[[9]], files[[10]], files[[11]], files[[12]]
    )
  )
  seurat$sample <- c(
    rep(sub("\\/.*", "", file_list[1]), ncol(files[[1]])),
    rep(sub("\\/.*", "", file_list[2]), ncol(files[[2]])),
    rep(sub("\\/.*", "", file_list[3]), ncol(files[[3]])),
    rep(sub("\\/.*", "", file_list[4]), ncol(files[[4]])),
    rep(sub("\\/.*", "", file_list[5]), ncol(files[[5]])),
    rep(sub("\\/.*", "", file_list[6]), ncol(files[[6]])),
    rep(sub("\\/.*", "", file_list[7]), ncol(files[[7]])),
    rep(sub("\\/.*", "", file_list[8]), ncol(files[[8]])),
    rep(sub("\\/.*", "", file_list[9]), ncol(files[[9]])),
    rep(sub("\\/.*", "", file_list[10]), ncol(files[[10]])),
    rep(sub("\\/.*", "", file_list[11]), ncol(files[[11]])),
    rep(sub("\\/.*", "", file_list[12]), ncol(files[[12]]))
  )

  saveRDS(seurat, rds)
}
```

```{r}
DefaultAssay(seurat) <- "RNA"
seurat[["SCT"]] <- NULL
sub_name <- "all"
seurat <- cluster_pipeline(
  seurat, cache_dir, sub_name, protocol, vars_to_regress, parallel_override, FALSE
)

for (i in 1:length(metadata_to_plot)) {
  print(names(metadata_to_plot)[i])
  if ("no_legend" %in% metadata_to_plot[[i]] && "no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(seurat, "umap1", "umap2", names(metadata_to_plot)[i]) + NoLegend())
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat") + NoLegend())
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(seurat, "umap1", "umap2", names(metadata_to_plot)[i]))
  } else {
    print(red_dim_plot(seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat"))
  }
}
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
