---
title: "02 Annotation - Combined Organelles"
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "Diana Acosta [Diana.Acosta@osumc.edu]"
  - name: "Hongjun Fu [Hongjun.Fu@osumc.edu]"
  - name: "Ohio State University Wexner Medical Center"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[8], "02_annotation.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Acosta Organoid scRNAseq Project](https://github.com/eturkes/acosta-organoid-scRNAseq).*

In this document we annotate cell-types based on reference datasets and marker genes.
The data here will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[8]``.

```{r}
#    This file is part of acosta-organoid-scRNAseq.
#    Copyright (C) 2022  Emir Turkes, Diana Acosta, Hongjun Fu, Ohio
#    State University Wexner Medical Center
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

packages <- c(
  "conflicted", "Seurat", "openxlsx", "stringr", "ggplot2", "patchwork", "GSVA", "GSEABase",
  "SingleCellExperiment", "scuttle", "SingleR", "dplyr", "ggrepel", "DT"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path("..", "utils.R"))

analysis_no <- 2
data_name <- unlist(strsplit(getwd(), "/"))[8] # Name of dataset.
data_dir <- file.path("..", "..", "data") # Backed up data.
results_dir <- file.path("..", "..", "results")

# Unique cache directory for each analysis number.
# ------------------------------------------------
cache_dir <- file.path("..", "..", "cache", data_name, paste0("0", analysis_no))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# ------------------------------------------------

knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
```

# Prep

```{r}
seurat <- readRDS(file.path(cache_dir, "..", "01", "all_seurat.rds"))
seurat
```

# Clustering and Marker Genes

```{r}
seurat <- FindClusters(seurat, verbose = FALSE)
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

rds <- file.path(cache_dir, "markers.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

datatable_download(markers)

markers %>%
    group_by(cluster) %>%
    top_n(n = 1, wt = avg_log2FC) -> top1

for (i in seq_along(top1$gene)) {
  print(FeaturePlot(seurat, features = top1$gene[i], order = TRUE, cols = c("lightgrey", "red")))
}
```

# SingleR

```{r}
load(file.path(data_dir, "polioudakis_2019", "raw_counts_mat.rdata"))
cell_metadata <- read.csv(file.path(data_dir, "polioudakis_2019", "cell_metadata.csv"))
raw_counts_mat <- raw_counts_mat[ , colnames(raw_counts_mat) %in% cell_metadata$Cell]

ref <- SingleCellExperiment(list(counts = raw_counts_mat), colData = cell_metadata)
ref <- logNormCounts(ref)

rds <- file.path(cache_dir, "singleR_clusters.rds")
if (file.exists(rds)) {
  predictions <- readRDS(rds)
} else {
  predictions <- SingleR(
    GetAssayData(seurat, assay = "RNA"), ref, ref$Cluster, clusters = seurat$seurat_clusters
  )
  saveRDS(predictions, rds)
}

predictions <- predictions$labels
names(predictions) <- levels(seurat)
seurat <- RenameIdents(seurat, predictions)
seurat$polioudakis_singleR_clusters <- seurat@active.ident

red_dim_plot(seurat, "umap1", "umap2", "polioudakis_singleR_clusters", "cat")
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
