---
title: "02 Annotation - Combined Organelles"
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "Diana Acosta [Diana.Acosta@osumc.edu]"
  - name: "Hongjun Fu [Hongjun.Fu@osumc.edu]"
  - name: "Ohio State University Wexner Medical Center"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[8], "02_annotation.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Acosta Organoid scRNAseq Project](https://github.com/eturkes/acosta-organoid-scRNAseq).*

In this document we annotate cell-types based on reference datasets and marker genes.
The data here will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[8]``.

```{r}
#    This file is part of acosta-organoid-scRNAseq.
#    Copyright (C) 2022  Emir Turkes, Diana Acosta, Hongjun Fu, Ohio
#    State University Wexner Medical Center
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

packages <- c(
  "conflicted", "Seurat", "openxlsx", "stringr", "ggplot2", "patchwork", "GSVA", "GSEABase",
  "SingleCellExperiment", "scuttle", "SingleR", "dplyr", "ggrepel", "DT", "edgeR", "ComplexHeatmap",
  "biomaRt"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
`%notin%` <- Negate(`%in%`)
source(file.path("..", "utils.R"))

analysis_no <- 2
data_name <- unlist(strsplit(getwd(), "/"))[8] # Name of dataset.
data_dir <- file.path("..", "..", "data") # Backed up data.
results_dir <- file.path("..", "..", "results")

# Unique cache directory for each analysis number.
# ------------------------------------------------
cache_dir <- file.path("..", "..", "cache", data_name, paste0("0", analysis_no))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# ------------------------------------------------

knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
```

# Prep

```{r}
rds <- file.path(cache_dir, "reclustered_seurat.rds")
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- readRDS(file.path(cache_dir, "..", "01", "all_seurat.rds"))
  seurat <- RunUMAP(seurat, 1:9, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2
  seurat <- FindNeighbors(seurat, dims = 1:9, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = 0.1, verbose = FALSE)
  saveRDS(seurat, rds)
}

seurat
```

# Clustering and Marker Genes

```{r}
ElbowPlot(seurat, 50) + geom_vline(xintercept = 9, color = "red")
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

rds <- file.path(cache_dir, "markers.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

datatable_download(markers)

markers %>%
    group_by(cluster) %>%
    top_n(n = 1, wt = avg_log2FC) -> top

for (i in seq_along(top$gene)) {
  print(FeaturePlot(seurat, features = top$gene[i], order = TRUE, cols = c("lightgrey", "red")))
}
```

# SingleR

## Polioudakis on Clusters

```{r}
rds <- file.path(cache_dir, "singleR_clusters.rds")
if (file.exists(rds)) {
  predictions <- readRDS(rds)
} else {

  load(file.path(data_dir, "polioudakis_2019", "raw_counts_mat.rdata"))
  cell_metadata <- read.csv(file.path(data_dir, "polioudakis_2019", "cell_metadata.csv"))
  raw_counts_mat <- raw_counts_mat[ , colnames(raw_counts_mat) %in% cell_metadata$Cell]
  
  ref <- SingleCellExperiment(list(counts = raw_counts_mat), colData = cell_metadata)
  ref <- logNormCounts(ref)

  predictions <- SingleR(
    GetAssayData(seurat, assay = "RNA"), ref, ref$Cluster, clusters = seurat$seurat_clusters
  )

  saveRDS(predictions, rds)
}

predictions <- predictions$labels
names(predictions) <- levels(seurat)
seurat <- RenameIdents(seurat, predictions)
seurat$polioudakis_singleR_clusters <- seurat@active.ident

red_dim_plot(seurat, "umap1", "umap2", "polioudakis_singleR_clusters", "cat")
```

# GSVA

```{r}
rds <- file.path(cache_dir, "pseudobulk.rds")
if (file.exists(rds)) {
  pseudobulk <- readRDS(rds)
} else {
  pseudobulk <- as.SingleCellExperiment(seurat)
  counts(pseudobulk) <- logcounts(pseudobulk)
  pseudobulk <- suppressWarnings(aggregateAcrossCells(pseudobulk, seurat$seurat_clusters))
  dge <- DGEList(counts(pseudobulk))
  dge <- calcNormFactors(dge)
  logcounts(pseudobulk, withDimnames = FALSE) <- edgeR::cpm(dge, log = TRUE, prior.count = 3)
  saveRDS(pseudobulk, rds)
}
```

## Polioudakis

```{r}
markers <- getGmt(file.path(data_dir, "polioudakis_2019", "polioudakis-top40.gmt"))
ast <- geneIds(markers[["Ast"]])
meta <- read.xlsx(
  file.path(data_dir, "polioudakis_2019", "1-s2.0-S0896627319305616-mmc5.xlsx"), 3
)
markers <- vector("list", length(unique(meta$Cluster)) + 1)
names(markers) <- c(unique(meta$Cluster), "Ast")
for (i in seq_along(markers)) {
  markers[[i]] <- meta$Gene[meta$Cluster %in% unique(meta$Cluster)[i]]
}
markers[[length(unique(meta$Cluster)) + 1]] <- ast

gsva <- gsva(
  logcounts(pseudobulk), markers, method = "ssgsea", ssgsea.norm = FALSE, verbose = FALSE
)
gsva <- (2 * (gsva - min(gsva)) / (max(gsva) - min(gsva))) - 1

set.seed(1)
heatmap <- Heatmap(
  gsva,
  column_km = 6,
  show_column_names = FALSE,
  cluster_columns = FALSE,
  row_names_max_width = max_text_width(rownames(gsva)),
  heatmap_legend_param = list(title = "SSGSEA enrichment", title_position = "leftcenter-rot"))
draw(heatmap)

FeaturePlot(seurat, markers[["IP"]][1:4])
```

## Polioudakis (No Duplicates)

```{r}
meta_extend <- rbind(meta, meta[1:length(ast), ])
meta_extend$Gene[(nrow(meta) + 1):(nrow(meta) + length(ast))] <- ast
meta_extend$Cluster[(nrow(meta) + 1):(nrow(meta) + length(ast))] <- "Ast"

dup <- unique(meta_extend$Gene[duplicated(meta_extend$Gene)])
meta_nodup <- meta_extend[-which(meta_extend$Gene %in% dup), ]

markers <- vector("list", length(unique(meta_nodup$Cluster)))
names(markers) <- c(unique(meta_nodup$Cluster))
for (i in seq_along(markers)) {
  markers[[i]] <- meta_nodup$Gene[meta_nodup$Cluster %in% unique(meta_nodup$Cluster)[i]]
}

gsva <- gsva(
  logcounts(pseudobulk), markers, method = "ssgsea", ssgsea.norm = FALSE, verbose = FALSE
)
gsva <- (2 * (gsva - min(gsva)) / (max(gsva) - min(gsva))) - 1

set.seed(1)
heatmap <- Heatmap(
  gsva,
  column_km = 6,
  show_column_names = FALSE,
  cluster_columns = FALSE,
  row_names_max_width = max_text_width(rownames(gsva)),
  heatmap_legend_param = list(title = "SSGSEA enrichment", title_position = "leftcenter-rot"))
draw(heatmap)
```

## Polioudakis (No Duplicates, All Ast)

```{r}
meta_extend <- rbind(meta, meta[1:length(ast), ])
meta_extend$Gene[(nrow(meta) + 1):(nrow(meta) + length(ast))] <- ast
meta_extend$Cluster[(nrow(meta) + 1):(nrow(meta) + length(ast))] <- "Ast"

dup <- unique(meta_extend$Gene[duplicated(meta_extend$Gene)])
meta_nodup <- meta_extend[-which(meta_extend$Gene %in% dup), ]

markers <- vector("list", length(unique(meta_nodup$Cluster)))
names(markers) <- c(unique(meta_nodup$Cluster))
for (i in seq_along(markers)) {
  markers[[i]] <- meta_nodup$Gene[meta_nodup$Cluster %in% unique(meta_nodup$Cluster)[i]]
}
markers[["Ast"]] <- ast

gsva <- gsva(
  logcounts(pseudobulk), markers, method = "ssgsea", ssgsea.norm = FALSE, verbose = FALSE
)
gsva <- (2 * (gsva - min(gsva)) / (max(gsva) - min(gsva))) - 1

set.seed(1)
heatmap <- Heatmap(
  gsva,
  column_km = 6,
  show_column_names = TRUE,
  cluster_columns = FALSE,
  row_names_max_width = max_text_width(rownames(gsva)),
  heatmap_legend_param = list(title = "SSGSEA enrichment", title_position = "leftcenter-rot"))
draw(heatmap)
```

## Bhaduri

```{r}
bhaduri_markers <- read.xlsx(
  file.path(data_dir, "bhaduri_2020", "41586_2020_1962_MOESM3_ESM.xlsx"), 2
)
bhaduri_markers <- bhaduri_markers[ , (ncol(bhaduri_markers) - 2):ncol(bhaduri_markers)]
bhaduri_markers <- bhaduri_markers[-which(is.na(bhaduri_markers$Type)), ]

bhaduri_markers <- bhaduri_markers[-which(bhaduri_markers$Subtype == "Low Quality"), ]
bhaduri_markers <- bhaduri_markers[-which(bhaduri_markers$Subtype == "Microglia low quality"), ]
bhaduri_markers <- bhaduri_markers[-which(bhaduri_markers$Subtype == "Outlier"), ]

bhaduri_markers$Rationale[bhaduri_markers$Subtype == "IPC/newborn"] <-
  "PPP1R17, SSTR2, KCNQ3, ROBO2"
bhaduri_markers$Rationale[bhaduri_markers$Subtype == "Endothelial"] <- "CD93, GBN, IL32, IGFPB7"

bhaduri_markers$TypeSubtype <- paste(bhaduri_markers$Type, bhaduri_markers$Subtype, sep = "_")
dup <- bhaduri_markers[duplicated(bhaduri_markers$TypeSubtype), ]
bhaduri_markers <- bhaduri_markers[-which(duplicated(bhaduri_markers$TypeSubtype)), ]
for (i in seq(nrow(dup))) {
  idx <- match(dup$TypeSubtype[i], bhaduri_markers$TypeSubtype)
  bhaduri_markers$Rationale[idx] <- paste(
    bhaduri_markers$Rationale[idx], dup$Rationale[i], sep = ", "
  )
}

missing_genes <- unlist(str_split(bhaduri_markers$Rationale, ", "))
missing_genes <- missing_genes[missing_genes %notin% rownames(seurat)]
missing_genes <- missing_genes[-which(missing_genes == "")]
expanded_genes <- vector("list", length(missing_genes))
for (i in seq_along(missing_genes)) {
  expanded_genes[[i]] <- rownames(seurat)[grep(missing_genes[i], rownames(seurat))]
}

add_genes <- vector("list", 4)
add_genes[[1]] <- expanded_genes[[1]][-1]
add_genes[[2]] <- expanded_genes[[2]]
add_genes[[3]] <- expanded_genes[[4]]
add_genes[[4]] <- expanded_genes[[6]][-1]
names(add_genes) <- c("AURK", "HMGB", "ROBO", "AIF1")

replace_idx <- vector("list", 4)
for (i in seq_along(replace_idx)) {
  replace_idx[[i]] <- grep(names(add_genes)[i], str_split(bhaduri_markers$Rationale, ", "))
}

bhaduri_markers$Rationale[1] <-
  "TOP2A, AURKB, AURKC, AURKA, HMGB2, HMGB1, HMGB3, CTNNB1, HES1, VIM, PTN, CLU"
bhaduri_markers$Rationale[5] <- "MEF2C, STMN2, FLT, ROBO2, ROBO1, ROBO3, CRYM"
bhaduri_markers$Rationale[7] <- "C1QB, AIF1L, CCL4, C1QC, AIF1, SPP1, C1QC, CCL3"

markers <- str_split(bhaduri_markers$Rationale, ", ")
names(markers) <- bhaduri_markers$TypeSubtype

gsva <- gsva(
  logcounts(pseudobulk), markers, method = "ssgsea", ssgsea.norm = FALSE, verbose = FALSE
)
gsva <- (2 * (gsva - min(gsva)) / (max(gsva) - min(gsva))) - 1

set.seed(1)
heatmap <- Heatmap(
  gsva,
  column_km = 6,
  show_column_names = TRUE,
  cluster_columns = FALSE,
  row_names_max_width = max_text_width(rownames(gsva)),
  heatmap_legend_param = list(title = "SSGSEA enrichment", title_position = "leftcenter-rot"))
draw(heatmap)
```

# Custom List

```{r}
markers <- read.delim(
  file.path(data_dir, "gene-lists", "Human cortical development-gene-list summary - page3.tsv"),
  check.names = FALSE
)

name_idx <- which(colnames(markers) == "Pollen Cell 2015")
start_idx <- name_idx + 1
end_idx <- which(colnames(markers) == "Bhaduri-Nature 2020") - 1
colnames(markers)[start_idx:end_idx] <- paste(
  colnames(markers)[name_idx], colnames(markers)[start_idx:end_idx], sep = "_"
)

name_idx <- which(colnames(markers) == "Bhaduri-Nature 2020")
start_idx <- name_idx + 1
end_idx <- which(colnames(markers) == "Polioudakis Neuron 2019") - 1
colnames(markers)[start_idx:end_idx] <- paste(
  colnames(markers)[name_idx], colnames(markers)[start_idx:end_idx], sep = "_"
)

name_idx <- which(colnames(markers) == "Polioudakis Neuron 2019")
start_idx <- name_idx + 1
end_idx <- ncol(markers)
colnames(markers)[start_idx:end_idx] <- paste(
  colnames(markers)[name_idx], colnames(markers)[start_idx:end_idx], sep = "_"
)

markers <- Filter(function(x) !all(is.na(x)), markers)
markers <- as.list(markers)
markers <- lapply(markers, function(x) x[!(x == "")])

rds <- file.path(cache_dir, "gene_anno.rds")
if (file.exists(rds)) {
  gene_anno <- readRDS(rds)
} else {
  ensembl <- unique(unlist(markers)[grep("ENS", unlist(markers))])
  mart <- useEnsembl("ensembl", "hsapiens_gene_ensembl")
  attributes <- c("external_gene_name", "ensembl_gene_id")
  gene_anno <- getBM(attributes, "ensembl_gene_id", ensembl, mart)
  saveRDS(gene_anno, rds)
}

gene_anno <- gene_anno[-which(gene_anno$external_gene_name == ""), ]
remove <- ensembl[ensembl %notin% gene_anno$ensembl_gene_id]
markers <- lapply(markers, function(x) x[!(x %in% remove)])
for (i in seq(nrow(gene_anno))) {
  markers <- rapply(
    markers,
    function(x) ifelse(x == gene_anno$ensembl_gene_id[i], gene_anno$external_gene_name[i], x),
    how = "replace")
}

date_genes <- read.delim(file.path(data_dir, "gene-lists", "date-genes.tsv"))
date_genes <- date_genes[date_genes$Exc_Genes %in% unlist(markers), ]
for (i in seq(nrow(date_genes))) {
  markers <- rapply(
    markers,
    function(x) ifelse(x == date_genes$Exc_Genes[i], date_genes$Gene_Symbol[i], x),
    how = "replace")
}

gsva <- gsva(
  logcounts(pseudobulk), markers, method = "ssgsea", ssgsea.norm = FALSE, verbose = FALSE
)
gsva <- (2 * (gsva - min(gsva)) / (max(gsva) - min(gsva))) - 1

set.seed(1)
heatmap <- Heatmap(
  gsva,
  column_km = 6,
  show_column_names = TRUE,
  cluster_columns = FALSE,
  row_names_max_width = max_text_width(rownames(gsva)),
  heatmap_legend_param = list(title = "SSGSEA enrichment", title_position = "leftcenter-rot"))
draw(heatmap)
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
