---
title: '02 Annotation - `r unlist(strsplit(getwd(), "/"))[8]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "Diana Acosta [Diana.Acosta@osumc.edu]"
  - name: "Hongjun Fu [Hongjun.Fu@osumc.edu]"
  - name: "Ohio State University Wexner Medical Center"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[8], "02_annotation.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Acosta Organoid scRNAseq Project](https://github.com/eturkes/acosta-organoid-scRNAseq).*

The data here will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[8]``.

```{r}
#    This file is part of acosta-organoid-scRNAseq.
#    Copyright (C) 2022  Emir Turkes, Diana Acosta, Hongjun Fu, Ohio
#    State University Wexner Medical Center
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

packages <- c(
  "conflicted", "Seurat", "data.table", "SingleCellExperiment", "Matrix", "scater", "SingleR",
  "dplyr", "ggrepel"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path("..", "utils.R"))

analysis_no <- 2
data_name <- unlist(strsplit(getwd(), "/"))[8] # Name of dataset.
data_dir <- file.path("..", "..", "data") # Backed up data.
results_dir <- file.path("..", "..", "results")

# Unique cache directory for each analysis number.
# ------------------------------------------------
cache_dir <- file.path("..", "..", "cache", data_name, paste0("0", analysis_no))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# ------------------------------------------------

knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
```

# Prep

```{r}
seurat <- readRDS(file.path(cache_dir, "..", "01", "all_seurat.rds"))
seurat

rds <- file.path(cache_dir, "data_list.rds")
if (file.exists(rds)) {
  data_list <- readRDS(rds)
} else {

  data_list <- vector("list", 4)

  data <- fread(file.path(data_dir, "bhaduri_2020", "exprMatrix.tsv"), data.table = FALSE)
  rownames(data) <- gsub(".+[|]", "", data$gene)
  data <- data[ , -1]
  meta <- read.delim(file.path(data_dir, "bhaduri_2020", "meta.tsv"))
  meta$TypeSubtype <- paste(meta$Type, meta$Subtype, sep = "_")
  data_list[[1]] <- SingleCellExperiment(list(logcounts = data), colData = meta)

  load(file.path(data_dir, "polioudakis_2019", "raw_counts_mat.rdata"))
  meta <- read.csv(file.path(data_dir, "polioudakis_2019", "cell_metadata.csv"))
  raw_counts_mat <- raw_counts_mat[ , colnames(raw_counts_mat) %in% meta$Cell]
  data_list[[2]] <- SingleCellExperiment(list(counts = raw_counts_mat), colData = meta)
  rm(raw_counts_mat)
  data_list[[2]] <- logNormCounts(data_list[[2]])

  data <- readMM(file.path(data_dir, "kanton_2019", "human_cell_counts_GRCh38.mtx"))
  meta <- read.delim(file.path(data_dir, "kanton_2019", "metadata_human_cells.tsv"))
  genes <- read.delim(file.path(data_dir, "kanton_2019", "genes_GRCh38.txt"), header = FALSE)
  colnames(data) <- rownames(meta)
  rownames(data) <- genes$V2
  data_list[[3]] <- SingleCellExperiment(list(counts = data), colData = meta, rowData = genes)
  rm(genes)
  data_list[[3]] <- logNormCounts(data_list[[3]])

  file_list <- list.files(file.path(data_dir, "velasco_2019", "SCP282", "expression"))
  for (i in seq_along(file_list)) {
    if (i == 1) {
      data <- fread(
        file.path(data_dir, "velasco_2019", "SCP282", "expression", file_list[i]),
        data.table = FALSE
      )
    } else {
      data <- cbind(
        data,
        fread(
          file.path(data_dir, "velasco_2019", "SCP282", "expression", file_list[i]),
          data.table = FALSE
        )
      )
    }
  }
  meta <- read.delim(file.path(data_dir, "velasco_2019", "SCP282", "metadata", "meta_combined.txt"))
  meta <- meta[-1, ]
  data <- data[ , colnames(data) %in% meta$NAME]
  data <- data[ , order(match(colnames(data), meta$NAME))]
  data_list[[4]] <- SingleCellExperiment(list(logcounts = data), colData = meta)
  
  rm(data, meta)
  
  saveRDS(data_list, rds)
}

for (i in seq_along(data_list)) {
  print(data_list[[i]])
}
```

# Annotation

```{r}
rds <- file.path(cache_dir, "predictions.rds")
if (file.exists(rds)) {
  predictions <- readRDS(rds)
} else {
  predictions <- SingleR(
    GetAssayData(seurat, assay = "RNA"),
    list(
      bhaduri = data_list[[1]], polioudakis = data_list[[2]],
      kanton_2019 = data_list[[3]], velasco = data_list[[4]]
    ),
    list(
      data_list[[1]]$TypeSubtype, data_list[[2]]$Subcluster,
      data_list[[3]]$PredCellType, data_list[[4]]$CellType
    ),
    aggr.ref = TRUE
  )
  saveRDS(predictions, rds)
}

table(predictions$labels)
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
