---
title: '02 Annotation - `r unlist(strsplit(getwd(), "/"))[8]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "Diana Acosta [Diana.Acosta@osumc.edu]"
  - name: "Hongjun Fu [Hongjun.Fu@osumc.edu]"
  - name: "Ohio State University Wexner Medical Center"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[8], "02_annotation.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Acosta Organoid scRNAseq Project](https://github.com/eturkes/acosta-organoid-scRNAseq).*

The data here will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[8]``.

```{r}
#    This file is part of acosta-organoid-scRNAseq.
#    Copyright (C) 2022  Emir Turkes, Diana Acosta, Hongjun Fu, Ohio
#    State University Wexner Medical Center
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

packages <- c(
  "conflicted", "Seurat", "openxlsx", "stringr", "ggplot2", "patchwork", "GSEABase", "GSVA",
  "dplyr", "ggrepel", "ComplexHeatmap"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path("..", "utils.R"))

analysis_no <- 2
data_name <- unlist(strsplit(getwd(), "/"))[8] # Name of dataset.
data_dir <- file.path("..", "..", "data") # Backed up data.
results_dir <- file.path("..", "..", "results")

# Unique cache directory for each analysis number.
# ------------------------------------------------
cache_dir <- file.path("..", "..", "cache", data_name, paste0("0", analysis_no))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# ------------------------------------------------

knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
```

# Prep

```{r}
seurat <- readRDS(file.path(cache_dir, "..", "01", "all_seurat.rds"))
seurat

bhaduri_markers <- read.xlsx(
  file.path(data_dir, "bhaduri_2020", "41586_2020_1962_MOESM3_ESM.xlsx"), 2
)
bhaduri_markers <- bhaduri_markers[ , (ncol(bhaduri_markers) - 2):ncol(bhaduri_markers)]
bhaduri_markers <- bhaduri_markers[-which(is.na(bhaduri_markers$Type)), ]
bhaduri_markers <- bhaduri_markers[-which(bhaduri_markers$Subtype == "Low Quality"), ]
bhaduri_markers <- bhaduri_markers[-which(bhaduri_markers$Subtype == "Microglia low quality"), ]
bhaduri_markers <- bhaduri_markers[-which(bhaduri_markers$Subtype == "Outlier"), ]
bhaduri_markers$Rationale[bhaduri_markers$Subtype == "IPC/newborn"] <-
  c("PPP1R17, SSTR2, KCNQ3, ROBO2")
```

# Bhaduri et al. 2020 Marker Genes

```{r}
for (i in seq(nrow(bhaduri_markers))) {
  genes <- str_split(bhaduri_markers$Rationale[i], ", ")[[1]]

  if (length(which(genes %in% rownames(seurat))) > 0) {

    if (length(which(genes %in% rownames(seurat))) > 4) {
      ncol <- 3
    } else {
      ncol <- 2
    }

    print(
      FeaturePlot(
        seurat, genes, order = TRUE, cols = c("lightgrey", "red"),
        ncol = ncol, keep.scale = "all", reduction = "umap"
      ) +
        plot_annotation(
          paste(bhaduri_markers$Type[i], bhaduri_markers$Subtype[i], sep = " - "),
          theme = theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"))
        )
    )
  }
}
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
