---
title: "02 Annotation - Combined Organelles"
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "Diana Acosta [Diana.Acosta@osumc.edu]"
  - name: "Hongjun Fu [Hongjun.Fu@osumc.edu]"
  - name: "Ohio State University Wexner Medical Center"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[8], "02_annotation.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Acosta Organoid scRNAseq Project](https://github.com/eturkes/acosta-organoid-scRNAseq).*

In this document we annotate cell-types based on reference datasets and marker genes.
The data here will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[8]``.

```{r}
#    This file is part of acosta-organoid-scRNAseq.
#    Copyright (C) 2022  Emir Turkes, Diana Acosta, Hongjun Fu, Ohio
#    State University Wexner Medical Center
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

packages <- c(
  "conflicted", "Seurat", "openxlsx", "stringr", "ggplot2", "patchwork", "GSVA", "GSEABase",
  "SingleCellExperiment", "scuttle", "SingleR", "dplyr", "ggrepel", "DT", "edgeR", "ComplexHeatmap",
  "biomaRt"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
`%notin%` <- Negate(`%in%`)
source(file.path("..", "utils.R"))

analysis_no <- 2
data_name <- unlist(strsplit(getwd(), "/"))[8] # Name of dataset.
data_dir <- file.path("..", "..", "data") # Backed up data.
results_dir <- file.path("..", "..", "results")

# Unique cache directory for each analysis number.
# ------------------------------------------------
cache_dir <- file.path("..", "..", "cache", data_name, paste0("0", analysis_no))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# ------------------------------------------------

knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
```

# Prep

```{r}
rds <- file.path(cache_dir, "reclustered_seurat.rds")
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- readRDS(file.path(cache_dir, "..", "01", "all_seurat.rds"))
  seurat <- RunUMAP(seurat, 1:39, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2
  seurat <- FindNeighbors(seurat, dims = 1:39, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = 1, verbose = FALSE)
  saveRDS(seurat, rds)
}

seurat
```

# Clustering and Marker Genes

```{r}
ElbowPlot(seurat, 50) + geom_vline(xintercept = 39, color = "red")
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat") + ggtitle("Dims = 39, Res = 1")
red_dim_plot(seurat, "umap1", "umap2", "sample", "cat") + ggtitle("Dims = 39, Res = 1")

rds <- file.path(cache_dir, "markers.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

datatable_download(markers)

markers %>%
    group_by(cluster) %>%
    top_n(n = 1, wt = avg_log2FC) -> top

for (i in seq_along(top$gene)) {
  print(FeaturePlot(seurat, features = top$gene[i], order = TRUE, cols = c("lightgrey", "red")))
}
```

# GSVA

```{r}
rds <- file.path(cache_dir, "pseudobulk.rds")
if (file.exists(rds)) {
  pseudobulk <- readRDS(rds)
} else {
  pseudobulk <- as.SingleCellExperiment(seurat)
  counts(pseudobulk) <- logcounts(pseudobulk)
  pseudobulk <- suppressWarnings(aggregateAcrossCells(pseudobulk, seurat$seurat_clusters))
  dge <- DGEList(counts(pseudobulk))
  dge <- calcNormFactors(dge)
  logcounts(pseudobulk, withDimnames = FALSE) <- edgeR::cpm(dge, log = TRUE, prior.count = 3)
  saveRDS(pseudobulk, rds)
}
```

## Bhaduri + Pollen

```{r}
markers <- read.delim(
  file.path(data_dir, "gene-lists", "Human cortical development-gene-list summary - page3.tsv"),
  check.names = FALSE
)

markers <- markers[ , 1:which(colnames(markers) == "Polioudakis Neuron 2019")]
markers <- markers[ , -(which(colnames(markers) == "oRG vRG shared genes"))]
markers <- Filter(function(x) !all(is.na(x)), markers)
markers <- as.list(markers)
markers <- lapply(markers, function(x) x[!(x == "")])

gsva <- gsva(
  logcounts(pseudobulk), markers, method = "ssgsea", ssgsea.norm = FALSE, verbose = FALSE
)
gsva <- (2 * (gsva - min(gsva)) / (max(gsva) - min(gsva))) - 1

set.seed(1)
heatmap <- Heatmap(
  gsva,
  column_km = 12,
  cluster_columns = FALSE,
  row_names_max_width = max_text_width(rownames(gsva)),
  heatmap_legend_param = list(title = "SSGSEA enrichment", title_position = "leftcenter-rot"))
draw(heatmap)

remove <- c(
  "SORCS1", "ADRA2A", "SLA", "STMN2", "SORBS2", "ELAVL2", "DCX", "ZFHX4", "GLI3", "CD99", "IGFBP7",
  "AURKC", "HMGB1", "HMGB3", "CTNNB1", "HES1", "VIM", "PTN", "CLU", "IL33", "NPR3", "CCNB2",
  "KIF15", "ARHGAP11B", "CENPK", "NEUROD2", "NEUROD6"
)
markers_trimmed <- markers
for (i in seq_along(remove)) {
  markers_trimmed <- rapply(
    markers_trimmed, function(x) ifelse(x == remove[i], "", x), how = "replace")
}
markers_trimmed <- lapply(markers_trimmed, function(x) x[!(x == "")])
remove <- which(
  names(markers_trimmed) %in% c(
    "Radial Glia_vRG", "Radial Glia_tRG", "Microglia", "oRG specific genes", "vRG specific genes"
  )
)
markers_trimmed <- markers_trimmed[-remove]

gsva <- gsva(
  logcounts(pseudobulk), markers_trimmed, method = "ssgsea", ssgsea.norm = FALSE, verbose = FALSE
)
gsva <- (2 * (gsva - min(gsva)) / (max(gsva) - min(gsva))) - 1

set.seed(1)
heatmap <- Heatmap(
  gsva,
  column_km = 12,
  cluster_columns = FALSE,
  row_names_max_width = max_text_width(rownames(gsva)),
  heatmap_legend_param = list(title = "SSGSEA enrichment", title_position = "leftcenter-rot"))
draw(heatmap)

out_dir <- file.path(results_dir, data_name, "markers-QC-trimmed")
if (!dir.exists(out_dir)) {
  dir.create(out_dir, recursive = TRUE)
}
for (i in seq_along(markers_trimmed)) {
  if (any(markers_trimmed[[i]] %in% rownames(seurat))) {
    png(
      file.path(out_dir, paste0("bhaduri_", names(markers_trimmed)[i], ".png")), 10, 7,
      "in", res = 300
    )
    plot(FeaturePlot(seurat, markers_trimmed[[i]], order = TRUE, cols = c("lightgrey", "red")))
    dev.off()
  }
}
```

### Clustering Based on GSVA

```{r}
rds <- file.path(cache_dir, "gsva.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(
    as.matrix(GetAssayData(seurat)), markers, method = "ssgsea",
    ssgsea.norm = FALSE, verbose = FALSE
  )
  saveRDS(gsva, rds)
}
gsva <- (2 * (gsva - min(gsva)) / (max(gsva) - min(gsva))) - 1

gsva_seurat <- CreateAssayObject(data = gsva)
seurat[["GSVA"]] <- gsva_seurat
DefaultAssay(seurat) <- "GSVA"

for (i in seq(nrow(seurat))) {
  print(
    FeaturePlot(seurat, rownames(seurat)[i], order = TRUE, reduction = "umap") +
      scale_color_gradientn(colors = c("blue", "#F8F8F8", "red"),  limits = c(-1, 1))
  )
}

set.seed(1)
heatmap <- Heatmap(
  gsva,
  column_km = 12,
  show_column_names = FALSE,
  column_gap = unit(3, "mm"),
  cluster_columns = FALSE,
  cluster_rows = FALSE,
  row_names_max_width = max_text_width(rownames(gsva)),
  heatmap_legend_param = list(title = "SSGSEA enrichment", title_position = "leftcenter-rot"))
draw(heatmap)

DefaultAssay(seurat) <- "SCT"
seurat[["GSVA"]] <- NULL
rds <- file.path(cache_dir, "gsva_trimmed.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(
    as.matrix(GetAssayData(seurat)), markers_trimmed, method = "ssgsea",
    ssgsea.norm = FALSE, verbose = FALSE
  )
  saveRDS(gsva, rds)
}
gsva <- (2 * (gsva - min(gsva)) / (max(gsva) - min(gsva))) - 1

gsva_seurat <- CreateAssayObject(data = gsva)
seurat[["GSVA"]] <- gsva_seurat
DefaultAssay(seurat) <- "GSVA"

DefaultAssay(seurat) <- "SCT"
seurat <- RunUMAP(
  seurat, features = rownames(seurat)[rownames(seurat) %in% unlist(markers_trimmed)],
  min.dist = 0.5, verbose = FALSE,
  reduction.name = "gsva_genes_umap", reduction.key = "GSVA_GENES_UMAP_"
)
add_df <- data.frame(Embeddings(seurat, "gsva_genes_umap"))
names(add_df) <- paste0("gsva_genes_umap", seq(ncol(add_df)))
seurat$gsva_genes_umap1 <- add_df$gsva_genes_umap1
seurat$gsva_genes_umap2 <- add_df$gsva_genes_umap2

png(
  file.path(out_dir, "art.png"), 10, 7,
  "in", res = 300
)
plot(red_dim_plot(seurat, "gsva_genes_umap1", "gsva_genes_umap2", "gsva_names") + NoLegend())
dev.off()

set.seed(1)
heatmap <- Heatmap(
  gsva,
  column_km = 19,
  show_column_names = FALSE,
  column_gap = unit(3, "mm"),
  cluster_columns = FALSE,
  cluster_rows = FALSE,
  row_names_max_width = max_text_width(rownames(gsva)),
  heatmap_legend_param = list(title = "SSGSEA enrichment", title_position = "leftcenter-rot"))
set.seed(1)
modules <- column_order(draw(heatmap))

mat <- GetAssayData(seurat, assay = "GSVA", slot = "data")
mat <- mat[ , unlist(modules)]
gsva_clusters <- data.frame(
  sample = colnames(mat),
  clusters = c(
    rep(names(lengths(modules))[1], lengths(modules)[1]),
    rep(names(lengths(modules))[2], lengths(modules)[2]),
    rep(names(lengths(modules))[3], lengths(modules)[3]),
    rep(names(lengths(modules))[4], lengths(modules)[4]),
    rep(names(lengths(modules))[5], lengths(modules)[5]),
    rep(names(lengths(modules))[6], lengths(modules)[6]),
    rep(names(lengths(modules))[7], lengths(modules)[7]),
    rep(names(lengths(modules))[8], lengths(modules)[8]),
    rep(names(lengths(modules))[9], lengths(modules)[9]),
    rep(names(lengths(modules))[10], lengths(modules)[10]),
    rep(names(lengths(modules))[11], lengths(modules)[11]),
    rep(names(lengths(modules))[12], lengths(modules)[12]),
    rep(names(lengths(modules))[13], lengths(modules)[13]),
    rep(names(lengths(modules))[14], lengths(modules)[14]),
    rep(names(lengths(modules))[15], lengths(modules)[15]),
    rep(names(lengths(modules))[16], lengths(modules)[16]),
    rep(names(lengths(modules))[17], lengths(modules)[17]),
    rep(names(lengths(modules))[18], lengths(modules)[18]),
    rep(names(lengths(modules))[19], lengths(modules)[19])
  )
)
gsva_clusters$seurat <- match(colnames(mat), colnames(seurat))
gsva_clusters <- gsva_clusters[order(gsva_clusters$seurat), ]
seurat$gsva_clusters <- gsva_clusters$clusters

seurat@active.ident <- factor(seurat$gsva_clusters)
seurat <- RenameIdents(
  seurat,
  "1" = "Inhibitory Neurons", "2" = "Excitatory Neuron Upper Layer",
  "3" = "Inhibitory Neurons", "4" = "Radial Glia Late",
  "5" = "Excitatory Neuron Newborn", "6" = "Unclassified",
  "7" = "Unclassified", "8" = "IPCs",
  "9" = "Excitatory Neuron Upper Layer", "10" = "Excitatory Neuron Upper Layer",
  "11" = "Excitatory Neuron Deep Layer", "12" = "Inhibitory Neurons",
  "13" = "Radial Glia Late", "14" = "Excitatory Neuron Upper Layer",
  "15" = "OPCs", "16" = "Mural Cells",
  "17" = "Radial Glia Early", "18" = "Radial Glia Early",
  "19" = "Astrocytes"
)
seurat$gsva_names <- seurat@active.ident
red_dim_plot(seurat, "umap1", "umap2", "gsva_names", "cat")

table(seurat$gsva_names, seurat$seurat_clusters)

seurat@active.ident <- seurat$seurat_clusters
seurat <- RenameIdents(
  seurat,
  "0" = "IPCs", "1" = "Unclassified",
  "2" = "Radial Glia Late", "3" = "Excitatory Neuron Deep Layer",
  "4" = "Unclassified", "5" = "Excitatory Neuron Upper Layer",
  "6" = "Excitatory Neuron Upper Layer", "7" = "Unclassified",
  "8" = "Radial Glia Early", "9" = "Excitatory Neuron Deep Layer",
  "10" = "Inhibitory Neurons", "11" = "Astrocytes",
  "12" = "Excitatory Neuron Newborn", "13" = "Radial Glia Late",
  "14" = "Radial Glia Early", "15" = "OPCs",
  "16" = "Inhibitory Neurons", "17" = "Radial Glia Late",
  "18" = "IPCs", "19" = "IPCs",
  "20" = "Unclassified", "21" = "Excitatory Neuron Upper Layer",
  "22" = "Radial Glia Late", "23" = "Mural Cells",
  "24" = "Radial Glia Late", "25" = "Excitatory Neuron Newborn",
  "26" = "Mural Cells", "27" = "OPCs",
  "28" = "Radial Glia Late", "29" = "Unclassified",
  "30" = "Radial Glia Early", "31" = "Radial Glia Early",
  "32" = "OPCs"
)
seurat$cluster_names <- seurat@active.ident
red_dim_plot(seurat, "umap1", "umap2", "cluster_names", "cat")
```

## Custom List

```{r}
markers <- read.delim(
  file.path(data_dir, "gene-lists", "Human cortical development-gene-list summary - page3.tsv"),
  check.names = FALSE
)

name_idx <- which(colnames(markers) == "Pollen Cell 2015")
start_idx <- name_idx + 1
end_idx <- which(colnames(markers) == "Bhaduri-Nature 2020") - 1
colnames(markers)[start_idx:end_idx] <- paste(
  colnames(markers)[name_idx], colnames(markers)[start_idx:end_idx], sep = "_"
)

name_idx <- which(colnames(markers) == "Bhaduri-Nature 2020")
start_idx <- name_idx + 1
end_idx <- which(colnames(markers) == "Polioudakis Neuron 2019") - 1
colnames(markers)[start_idx:end_idx] <- paste(
  colnames(markers)[name_idx], colnames(markers)[start_idx:end_idx], sep = "_"
)

name_idx <- which(colnames(markers) == "Polioudakis Neuron 2019")
start_idx <- name_idx + 1
end_idx <- which(colnames(markers) == "Kanton Nature 2019") - 1
colnames(markers)[start_idx:end_idx] <- paste(
  colnames(markers)[name_idx], colnames(markers)[start_idx:end_idx], sep = "_"
)

name_idx <- which(colnames(markers) == "Kanton Nature 2019")
start_idx <- name_idx + 1
end_idx <- ncol(markers)
colnames(markers)[start_idx:end_idx] <- paste(
  colnames(markers)[name_idx], colnames(markers)[start_idx:end_idx], sep = "_"
)

markers <- Filter(function(x) !all(is.na(x)), markers)
markers <- as.list(markers)
markers <- lapply(markers, function(x) x[!(x == "")])

ensembl <- unique(unlist(markers)[grep("ENS", unlist(markers))])
rds <- file.path(cache_dir, "gene_anno.rds")
if (file.exists(rds)) {
  gene_anno <- readRDS(rds)
} else {
  mart <- useEnsembl("ensembl", "hsapiens_gene_ensembl")
  attributes <- c("external_gene_name", "ensembl_gene_id")
  gene_anno <- getBM(attributes, "ensembl_gene_id", ensembl, mart)
  saveRDS(gene_anno, rds)
}

gene_anno <- gene_anno[-which(gene_anno$external_gene_name == ""), ]
remove <- ensembl[ensembl %notin% gene_anno$ensembl_gene_id]
markers <- lapply(markers, function(x) x[!(x %in% remove)])
for (i in seq(nrow(gene_anno))) {
  markers <- rapply(
    markers,
    function(x) ifelse(x == gene_anno$ensembl_gene_id[i], gene_anno$external_gene_name[i], x),
    how = "replace")
}

date_genes <- read.delim(file.path(data_dir, "gene-lists", "date-genes.tsv"))
date_genes <- date_genes[date_genes$Exc_Genes %in% unlist(markers), ]
for (i in seq(nrow(date_genes))) {
  markers <- rapply(
    markers,
    function(x) ifelse(x == date_genes$Exc_Genes[i], date_genes$Gene_Symbol[i], x),
    how = "replace")
}

gsva <- gsva(
  logcounts(pseudobulk), markers, method = "ssgsea", ssgsea.norm = FALSE, verbose = FALSE
)
gsva <- (2 * (gsva - min(gsva)) / (max(gsva) - min(gsva))) - 1

set.seed(1)
heatmap <- Heatmap(
  gsva,
  column_km = 12,
  cluster_columns = FALSE,
  cluster_rows = FALSE,
  row_names_max_width = max_text_width(rownames(gsva)),
  heatmap_legend_param = list(title = "SSGSEA enrichment", title_position = "leftcenter-rot"))
draw(heatmap)

gsva <- gsva(logcounts(pseudobulk), markers, method = "gsva", mx.diff = FALSE, verbose = FALSE)

set.seed(1)
heatmap <- Heatmap(
  gsva,
  column_km = 12,
  cluster_columns = FALSE,
  cluster_rows = FALSE,
  row_names_max_width = max_text_width(rownames(gsva)),
  heatmap_legend_param = list(title = "GSVA enrichment", title_position = "leftcenter-rot"))
draw(heatmap)
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
