---
title: "02 Annotation - Combined Organelles"
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "Diana Acosta [Diana.Acosta@osumc.edu]"
  - name: "Hongjun Fu [Hongjun.Fu@osumc.edu]"
  - name: "Ohio State University Wexner Medical Center"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[8], "02_annotation.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Acosta Organoid scRNAseq Project](https://github.com/eturkes/acosta-organoid-scRNAseq).*

In this document we annotate cell-types based on reference datasets and marker genes.
The data here will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[8]``.

```{r}
#    This file is part of acosta-organoid-scRNAseq.
#    Copyright (C) 2022  Emir Turkes, Diana Acosta, Hongjun Fu, Ohio
#    State University Wexner Medical Center
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

packages <- c(
  "conflicted", "Seurat", "ggplot2", "dplyr", "ggrepel", "SingleCellExperiment", "scuttle", "RCAv2",
  "data.table", "GSVA", "edgeR", "ComplexHeatmap"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path("..", "utils.R"))

analysis_no <- 2
data_name <- unlist(strsplit(getwd(), "/"))[8] # Name of dataset.
data_dir <- file.path("..", "..", "data") # Backed up data.
results_dir <- file.path("..", "..", "results")

# Unique cache directory for each analysis number.
# ------------------------------------------------
cache_dir <- file.path("..", "..", "cache", data_name, paste0("0", analysis_no))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# ------------------------------------------------

knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
```

# Prep

```{r}
rds <- file.path(cache_dir, "reclustered_seurat.rds")
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- readRDS(file.path(cache_dir, "..", "01", "all_seurat.rds"))
  seurat <- RunUMAP(seurat, 1:39, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2
  seurat <- FindNeighbors(seurat, dims = 1:39, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = 1, verbose = FALSE)
  saveRDS(seurat, rds)
}

seurat

ElbowPlot(seurat, 50) + geom_vline(xintercept = 39, color = "red")
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat") + ggtitle("Dims = 39, Res = 1")
red_dim_plot(seurat, "umap1", "umap2", "sample", "cat") + ggtitle("Dims = 39, Res = 1")
```

# ssGSEA + Unbiased Clustering

## Bhaduri

```{r}
markers <- read.delim(
  file.path(data_dir, "gene-lists", "Human cortical development-gene-list summary - page3.tsv"),
  check.names = FALSE
)

markers <- markers[ , 1:which(colnames(markers) == "Polioudakis Neuron 2019")]
markers <- markers[ , -(which(colnames(markers) == "oRG vRG shared genes"))]
markers <- Filter(function(x) !all(is.na(x)), markers)
markers <- as.list(markers)
markers <- lapply(markers, function(x) x[!(x == "")])

remove <- c(
  "SORCS1", "ADRA2A", "SLA", "STMN2", "SORBS2", "ELAVL2", "DCX", "ZFHX4", "GLI3", "CD99", "IGFBP7",
  "AURKC", "HMGB1", "HMGB3", "CTNNB1", "HES1", "VIM", "PTN", "CLU", "IL33", "NPR3", "CCNB2",
  "KIF15", "ARHGAP11B", "CENPK", "NEUROD2", "NEUROD6"
)
markers_trimmed <- markers
for (i in seq_along(remove)) {
  markers_trimmed <- rapply(
    markers_trimmed, function(x) ifelse(x == remove[i], "", x), how = "replace")
}
markers_trimmed <- lapply(markers_trimmed, function(x) x[!(x == "")])
remove <- which(
  names(markers_trimmed) %in% c(
    "Radial Glia_vRG", "Radial Glia_tRG", "Microglia", "oRG specific genes", "vRG specific genes"
  )
)
markers_trimmed <- markers_trimmed[-remove]

rds <- file.path(cache_dir, "seurat_pseudobulk.rds")
if (file.exists(rds)) {
  pseudobulk <- readRDS(rds)
} else {
  pseudobulk <- as.SingleCellExperiment(seurat)
  counts(pseudobulk) <- logcounts(pseudobulk)
  pseudobulk <- suppressWarnings(aggregateAcrossCells(pseudobulk, seurat$seurat_clusters))
  dge <- DGEList(counts(pseudobulk))
  dge <- calcNormFactors(dge)
  logcounts(pseudobulk, withDimnames = FALSE) <- edgeR::cpm(dge, log = TRUE, prior.count = 3)
  saveRDS(pseudobulk, rds)
}

gsva <- gsva(
  logcounts(pseudobulk), markers_trimmed, method = "ssgsea", ssgsea.norm = FALSE, verbose = FALSE
)
gsva <- (2 * (gsva - min(gsva)) / (max(gsva) - min(gsva))) - 1

set.seed(1)
heatmap <- Heatmap(
  gsva,
  column_km = 12,
  cluster_columns = FALSE,
  row_names_max_width = max_text_width(rownames(gsva)),
  heatmap_legend_param = list(title = "SSGSEA enrichment", title_position = "leftcenter-rot"))
draw(heatmap)

seurat@active.ident <- seurat$seurat_clusters
seurat <- RenameIdents(
  seurat,
  "0" = "IPC", "1" = "Unclassified 1",
  "2" = "Radial Glia Late", "3" = "Excitatory Neuron Deep Layer",
  "4" = "Unclassified 2", "5" = "Excitatory Neuron Upper Layer",
  "6" = "Excitatory Neuron Upper Layer", "7" = "Unclassified 3",
  "8" = "Radial Glia Early", "9" = "Excitatory Neuron Deep Layer",
  "10" = "Inhibitory Neurons", "11" = "Astrocyte",
  "12" = "Excitatory Neuron Newborn", "13" = "Radial Glia Late",
  "14" = "Radial Glia Early", "15" = "OPC",
  "16" = "Inhibitory Neurons", "17" = "Radial Glia Late",
  "18" = "IPC", "19" = "IPC",
  "20" = "Unclassified 4", "21" = "Excitatory Neuron Upper Layer",
  "22" = "Radial Glia Late", "23" = "Mural",
  "24" = "Radial Glia Late", "25" = "Excitatory Neuron Newborn",
  "26" = "Mural", "27" = "OPC",
  "28" = "Radial Glia Late", "29" = "Unclassified 3",
  "30" = "Radial Glia Early", "31" = "Radial Glia Early",
  "32" = "OPC"
)
seurat$cluster_names <- seurat@active.ident
red_dim_plot(seurat, "umap1", "umap2", "cluster_names", "cat")

for (i in seq_along(markers_trimmed)) {
  if (any(markers_trimmed[[i]] %in% rownames(seurat))) {
    print(names(markers_trimmed[i]))
    print(
      FeaturePlot(seurat, markers_trimmed[[i]], order = TRUE, cols = c("lightgrey", "red"))
    )
  }
}
```

# RCAv2

## Bhaduri

### Primary Broad Clusters

```{r}
rds <- file.path(cache_dir, "bhaduri_primary_rca_Type.rds")
rds2 <- file.path(cache_dir, "bhaduri_primary_pseudobulk_Type.rds")
if (file.exists(rds)) {
  rca <- readRDS(rds)
  ref <- readRDS(rds2)
} else {

  rds <- file.path(cache_dir, "bhaduri_primary_ref.rds")
  if (file.exists(rds)) {
    ref <- readRDS(rds)
  } else {
    ref <- fread(
      file.path(data_dir, "bhaduri_2020", "primary", "exprMatrix.tsv"), data.table = FALSE
    )
    gc()
    rownames(ref) <- ref[ , 1]
    ref <- ref[ , -1]
  
    meta <- read.delim(file.path(data_dir, "bhaduri_2020", "primary", "meta.tsv"), row.names = 1)
    meta <- meta[-(which(meta$Type == "Outlier")), ]
    meta <- meta[-(which(meta$Subtype == "Microglia low quality")), ]
    meta <- meta[-(which(meta$Subtype == "Low Quality")), ]
  
    ref <- ref[ , which(colnames(ref) %in% rownames(meta))]
    ref <- 2 ^ ref - 1
  
    ref <- SingleCellExperiment(list(counts = ref), colData = meta)
    rm(meta)
    gc()
    saveRDS(ref, rds)
  }

  rds <- file.path(cache_dir, "bhaduri_primary_pseudobulk_Type.rds")
  if (file.exists(rds)) {
    ref <- readRDS(rds)
  } else {
    top <- nrow(ref) / 2
    top <- ref[1:top, ]
    top <- aggregateAcrossCells(top, top$Type, "mean")
  
    bottom <- (nrow(ref) / 2) + 1
    bottom <- ref[bottom:(nrow(ref)), ]
    bottom <- aggregateAcrossCells(bottom, bottom$Type, "mean")
  
    ref <- rbind(top, bottom)
    rm(top, bottom)
    saveRDS(ref, rds)
  }

  rds <- file.path(cache_dir, "bhaduri_primary_panel_Type.rds")
  if (file.exists(rds)) {
    ref <- readRDS(rds)
  } else {
    ref <- ref[-(which(rowSums(counts(ref)) == 0)), ]
  
    ref2 <- ref
    colnames(ref2) <- paste(colnames(ref2), "2", sep = "_")
    colnames(ref) <- paste(colnames(ref), "1", sep = "_")
    ref <- cbind(ref, ref2)
    rm(ref2)
  
    ref <- buildReferencePanel(counts(ref))
    saveRDS(ref, rds)
  }

  rm(ref)
  rds <- file.path(cache_dir, "bhaduri_primary_rca_Type.rds")
  if (file.exists(rds)) {
    rca <- readRDS(rds)
  } else {
    rca <- createRCAObject(GetAssayData(seurat, "counts"), GetAssayData(seurat))
    rca <- dataProject(
      rca, "Custom", file.path(cache_dir, "bhaduri_primary_panel_Type.rds"), "spearman"
    )
    rca <- computeUMAP(rca)
    rca <- estimateCellTypeFromProjection(rca)
    saveRDS(rca, rds)
  }
}

rca

seurat$rca_clusters <- factor(unlist(rca$cell.Type.Estimate))
seurat[["rca_umap"]] <- CreateDimReducObject(
  as.matrix(rca$umap.coordinates), key = "rca_umap_", assay = DefaultAssay(seurat)
)
add_df <- data.frame(Embeddings(seurat, "rca_umap"))
names(add_df) <- paste0("rca_umap", seq(ncol(add_df)))
seurat$rca_umap1 <- add_df$rca_umap1
seurat$rca_umap2 <- add_df$rca_umap2

red_dim_plot(seurat, "rca_umap1", "rca_umap2", "rca_clusters", "cat")
red_dim_plot(seurat, "rca_umap1", "rca_umap2", "cluster_names", "cat")
red_dim_plot(seurat, "rca_umap1", "rca_umap2", "seurat_clusters", "cat")
red_dim_plot(seurat, "rca_umap1", "rca_umap2", "sample", "cat")

cat(
  length(unique(unlist(rca$cell.Type.Estimate))), "cell-types found:",
  unique(unlist(rca$cell.Type.Estimate))
)
cat("Out of", length(colnames(ref)), "possible:", colnames(ref))

for (i in seq_along(markers_trimmed)) {
  if (any(markers_trimmed[[i]] %in% rownames(seurat))) {
    print(names(markers_trimmed[i]))
    print(
      FeaturePlot(
        seurat, markers_trimmed[[i]], reduction = "rca_umap",
        order = TRUE, cols = c("lightgrey", "red")
      )
    )
  }
}
```

### Primary Sub Clusters

```{r}
rds <- file.path(cache_dir, "bhaduri_primary_rca_Subtype.rds")
rds2 <- file.path(cache_dir, "bhaduri_primary_pseudobulk_Subtype.rds")
if (file.exists(rds)) {
  rca <- readRDS(rds)
  ref <- readRDS(rds2)
} else {

  rds <- file.path(cache_dir, "bhaduri_primary_ref.rds")
  if (file.exists(rds)) {
    ref <- readRDS(rds)
  } else {
    ref <- fread(file.path(data_dir, "bhaduri_2020", "primary", "exprMatrix.tsv"), data.table = FALSE)
    gc()
    rownames(ref) <- ref[ , 1]
    ref <- ref[ , -1]
  
    meta <- read.delim(
      file.path(data_dir, "bhaduri_2020", "primary", "meta.tsv"), row.names = 1
    )
    meta <- meta[-(which(meta$Type == "Outlier")), ]
    meta <- meta[-(which(meta$Subtype == "Microglia low quality")), ]
    meta <- meta[-(which(meta$Subtype == "Low Quality")), ]
  
    ref <- ref[ , which(colnames(ref) %in% rownames(meta))]
  
    ref <- SingleCellExperiment(list(normcounts = ref), colData = meta)
    counts(ref) <- 2 ^ normcounts(ref) - 1
    rm(meta)
    gc()
    saveRDS(ref, rds)
  }

  rds <- file.path(cache_dir, "bhaduri_primary_pseudobulk_Subtype.rds")
  if (file.exists(rds)) {
    ref <- readRDS(rds)
  } else {
    top <- nrow(ref) / 2
    top <- ref[1:top, ]
    top <- aggregateAcrossCells(top, top$Subtype, "mean")
  
    bottom <- (nrow(ref) / 2) + 1
    bottom <- ref[bottom:(nrow(ref)), ]
    bottom <- aggregateAcrossCells(bottom, bottom$Subtype, "mean")
  
    ref <- rbind(top, bottom)
    rm(top, bottom)
    saveRDS(ref, rds)
  }

  rds <- file.path(cache_dir, "bhaduri_primary_panel_Subtype.rds")
  if (file.exists(rds)) {
    ref <- readRDS(rds)
  } else {
    ref <- ref[-(which(rowSums(counts(ref)) == 0)), ]
  
    ref2 <- ref
    colnames(ref2) <- paste(colnames(ref2), "2", sep = "_")
    colnames(ref) <- paste(colnames(ref), "1", sep = "_")
    ref <- cbind(ref, ref2)
    rm(ref2)
  
    ref <- buildReferencePanel(counts(ref))
    saveRDS(ref, rds)
  }

  rm(ref)
  rds <- file.path(cache_dir, "bhaduri_primary_rca_Subtype.rds")
  if (file.exists(rds)) {
    rca <- readRDS(rds)
  } else {
    rca <- createRCAObject(GetAssayData(seurat, "counts"), GetAssayData(seurat))
    rca <- dataProject(
      rca, "Custom", file.path(cache_dir, "bhaduri_primary_panel_Subtype.rds"), "spearman"
    )
    rca <- computeUMAP(rca)
    rca <- estimateCellTypeFromProjection(rca)
    saveRDS(rca, rds)
  }
}

rca

seurat$rca_clusters <- factor(unlist(rca$cell.Type.Estimate))
seurat[["rca_umap"]] <- CreateDimReducObject(
  as.matrix(rca$umap.coordinates), key = "rca_umap_", assay = DefaultAssay(seurat)
)
add_df <- data.frame(Embeddings(seurat, "rca_umap"))
names(add_df) <- paste0("rca_umap", seq(ncol(add_df)))
seurat$rca_umap1 <- add_df$rca_umap1
seurat$rca_umap2 <- add_df$rca_umap2

red_dim_plot(seurat, "rca_umap1", "rca_umap2", "rca_clusters", "cat")
red_dim_plot(seurat, "rca_umap1", "rca_umap2", "cluster_names", "cat")
red_dim_plot(seurat, "rca_umap1", "rca_umap2", "seurat_clusters", "cat")
red_dim_plot(seurat, "rca_umap1", "rca_umap2", "sample", "cat")

cat(
  length(unique(unlist(rca$cell.Type.Estimate))), "cell-types found:",
  unique(unlist(rca$cell.Type.Estimate))
)
cat("Out of", length(colnames(ref)), "possible:", colnames(ref))

for (i in seq_along(markers_trimmed)) {
  if (any(markers_trimmed[[i]] %in% rownames(seurat))) {
    print(names(markers_trimmed[i]))
    print(
      FeaturePlot(
        seurat, markers_trimmed[[i]], reduction = "rca_umap",
        order = TRUE, cols = c("lightgrey", "red")
      )
    )
  }
}
```

### Custom Clusters

```{r}
rds <- file.path(cache_dir, "bhaduri_primary_rca_custom.rds")
rds2 <- file.path(cache_dir, "bhaduri_primary_pseudobulk_custom.rds")
if (file.exists(rds)) {
  rca <- readRDS(rds)
  ref <- readRDS(rds2)
} else {

  rds <- file.path(cache_dir, "bhaduri_primary_ref.rds")
  if (file.exists(rds)) {
    ref <- readRDS(rds)
  } else {
    ref <- fread(file.path(data_dir, "bhaduri_2020", "primary", "exprMatrix.tsv"), data.table = FALSE)
    gc()
    rownames(ref) <- ref[ , 1]
    ref <- ref[ , -1]
  
    meta <- read.delim(
      file.path(data_dir, "bhaduri_2020", "primary", "meta.tsv"), row.names = 1
    )
    meta <- meta[-(which(meta$Type == "Outlier")), ]
    meta <- meta[-(which(meta$Subtype == "Microglia low quality")), ]
    meta <- meta[-(which(meta$Subtype == "Low Quality")), ]
  
    ref <- ref[ , which(colnames(ref) %in% rownames(meta))]
  
    ref <- SingleCellExperiment(list(normcounts = ref), colData = meta)
    counts(ref) <- 2 ^ normcounts(ref) - 1
    rm(meta)
    gc()
    saveRDS(ref, rds)
  }

  ref$custom <- ref$Subtype
  remove <- which(
    ref$custom == "oRG" | ref$custom == "vRG" | ref$custom == "IPC/newborn" |
      ref$custom == "Layer VI Occipital" | ref$custom == "Cajal Retzius" |
      ref$custom == "Layer VI Pan-area" | ref$custom == "PFC" |
      ref$custom == "Parietal and Temporal" | ref$custom == "Layer IV" | ref$custom == "tRG" |
      ref$custom == "Endothelial" | ref$custom == "Microglia"
  )
  ref <- ref[ , -remove]

  ref$custom[
    which(
      ref$custom == "Upper Layer Occipital" | ref$custom == "Upper Layer" |
        ref$custom == "Upper Layer PFC"
    )
  ] <- "Excitatory Neuron Upper Layer"
  ref$custom[
    which(ref$custom == "IPC-new" | ref$custom == "IPC-div1" | ref$custom == "IPC-div2")
  ] <- "IPC"
  ref$custom[which(ref$custom == "SST-MGE1" | ref$custom == "MGE2")] <- "Inhibitory Neuron"
  ref$custom[which(ref$custom == "Deep Layer")] <- "Excitatory Neuron Deep Layer"
  ref$custom[which(ref$custom == "late")] <- "Radial Glia Late"
  ref$custom[which(ref$custom == "early")] <- "Radial Glia Early"
  ref$custom[which(ref$custom == "Newborn")] <- "Excitatory Neuron Newborn"
  ref$custom[which(ref$custom == "oRG/Astrocyte")] <- "Astrocyte"

  rds <- file.path(cache_dir, "bhaduri_primary_pseudobulk_custom.rds")
  if (file.exists(rds)) {
    ref <- readRDS(rds)
  } else {
    top <- nrow(ref) / 2
    top <- ref[1:top, ]
    top <- aggregateAcrossCells(top, top$custom, "mean")
  
    bottom <- (nrow(ref) / 2) + 1
    bottom <- ref[bottom:(nrow(ref)), ]
    bottom <- aggregateAcrossCells(bottom, bottom$custom, "mean")
  
    ref <- rbind(top, bottom)
    rm(top, bottom)
    saveRDS(ref, rds)
  }

  rds <- file.path(cache_dir, "bhaduri_primary_panel_custom.rds")
  if (file.exists(rds)) {
    ref <- readRDS(rds)
  } else {
    ref <- ref[-(which(rowSums(counts(ref)) == 0)), ]
  
    ref2 <- ref
    colnames(ref2) <- paste(colnames(ref2), "2", sep = "_")
    colnames(ref) <- paste(colnames(ref), "1", sep = "_")
    ref <- cbind(ref, ref2)
    rm(ref2)
  
    ref <- buildReferencePanel(counts(ref))
    saveRDS(ref, rds)
  }

  rm(ref)
  rds <- file.path(cache_dir, "bhaduri_primary_rca_custom.rds")
  if (file.exists(rds)) {
    rca <- readRDS(rds)
  } else {
    rca <- createRCAObject(GetAssayData(seurat, "counts"), GetAssayData(seurat))
    rca <- dataProject(
      rca, "Custom", file.path(cache_dir, "bhaduri_primary_panel_custom.rds"), "spearman"
    )
    rca <- computeUMAP(rca)
    rca <- estimateCellTypeFromProjection(rca)
    saveRDS(rca, rds)
  }
}

rca

seurat$rca_clusters <- factor(unlist(rca$cell.Type.Estimate))
seurat[["rca_umap"]] <- CreateDimReducObject(
  as.matrix(rca$umap.coordinates), key = "rca_umap_", assay = DefaultAssay(seurat)
)
add_df <- data.frame(Embeddings(seurat, "rca_umap"))
names(add_df) <- paste0("rca_umap", seq(ncol(add_df)))
seurat$rca_umap1 <- add_df$rca_umap1
seurat$rca_umap2 <- add_df$rca_umap2

red_dim_plot(seurat, "rca_umap1", "rca_umap2", "rca_clusters", "cat")
red_dim_plot(seurat, "rca_umap1", "rca_umap2", "cluster_names", "cat")
red_dim_plot(seurat, "rca_umap1", "rca_umap2", "seurat_clusters", "cat")
red_dim_plot(seurat, "rca_umap1", "rca_umap2", "sample", "cat")

cat(
  length(unique(unlist(rca$cell.Type.Estimate))), "cell-types found:",
  unique(unlist(rca$cell.Type.Estimate))
)
cat("Out of", length(colnames(ref)), "possible:", colnames(ref))

for (i in seq_along(markers_trimmed)) {
  if (any(markers_trimmed[[i]] %in% rownames(seurat))) {
    print(names(markers_trimmed[i]))
    print(
      FeaturePlot(
        seurat, markers_trimmed[[i]], reduction = "rca_umap",
        order = TRUE, cols = c("lightgrey", "red")
      )
    )
  }
}
```

### Primary Broad Clusters Without Small Groups

```{r}
rds <- file.path(cache_dir, "bhaduri_primary_rca_Type_filt.rds")
rds2 <- file.path(cache_dir, "bhaduri_primary_pseudobulk_Type_filt.rds")
if (file.exists(rds)) {
  rca <- readRDS(rds)
  ref <- readRDS(rds2)
} else {

  rds <- file.path(cache_dir, "bhaduri_primary_ref.rds")
  if (file.exists(rds)) {
    ref <- readRDS(rds)
  } else {
    ref <- fread(
      file.path(data_dir, "bhaduri_2020", "primary", "exprMatrix.tsv"), data.table = FALSE
    )
    gc()
    rownames(ref) <- ref[ , 1]
    ref <- ref[ , -1]
  
    meta <- read.delim(file.path(data_dir, "bhaduri_2020", "primary", "meta.tsv"), row.names = 1)
    meta <- meta[-(which(meta$Type == "Outlier")), ]
    meta <- meta[-(which(meta$Subtype == "Microglia low quality")), ]
    meta <- meta[-(which(meta$Subtype == "Low Quality")), ]
  
    ref <- ref[ , which(colnames(ref) %in% rownames(meta))]
    ref <- 2 ^ ref - 1
  
    ref <- SingleCellExperiment(list(counts = ref), colData = meta)
    rm(meta)
    gc()
    saveRDS(ref, rds)
  }

  remove <- which(ref$Type == "OPC" | ref$Type == "Microglia" | ref$Type == "Endothelial")
  ref <- ref[ , -remove]

  rds <- file.path(cache_dir, "bhaduri_primary_pseudobulk_Type_filt.rds")
  if (file.exists(rds)) {
    ref <- readRDS(rds)
  } else {
    top <- nrow(ref) / 2
    top <- ref[1:top, ]
    top <- aggregateAcrossCells(top, top$Type, "mean")
  
    bottom <- (nrow(ref) / 2) + 1
    bottom <- ref[bottom:(nrow(ref)), ]
    bottom <- aggregateAcrossCells(bottom, bottom$Type, "mean")
  
    ref <- rbind(top, bottom)
    rm(top, bottom)
    saveRDS(ref, rds)
  }

  rds <- file.path(cache_dir, "bhaduri_primary_panel_Type_filt.rds")
  if (file.exists(rds)) {
    ref <- readRDS(rds)
  } else {
    ref <- ref[-(which(rowSums(counts(ref)) == 0)), ]
  
    ref2 <- ref
    colnames(ref2) <- paste(colnames(ref2), "2", sep = "_")
    colnames(ref) <- paste(colnames(ref), "1", sep = "_")
    ref <- cbind(ref, ref2)
    rm(ref2)
  
    ref <- buildReferencePanel(counts(ref))
    saveRDS(ref, rds)
  }

  rm(ref)
  rds <- file.path(cache_dir, "bhaduri_primary_rca_Type_filt.rds")
  if (file.exists(rds)) {
    rca <- readRDS(rds)
  } else {
    rca <- createRCAObject(GetAssayData(seurat, "counts"), GetAssayData(seurat))
    rca <- dataProject(
      rca, "Custom", file.path(cache_dir, "bhaduri_primary_panel_Type_filt.rds"), "spearman"
    )
    rca <- computeUMAP(rca)
    rca <- estimateCellTypeFromProjection(rca)
    saveRDS(rca, rds)
  }
}

rca

seurat$rca_clusters <- factor(unlist(rca$cell.Type.Estimate))
seurat[["rca_umap"]] <- CreateDimReducObject(
  as.matrix(rca$umap.coordinates), key = "rca_umap_", assay = DefaultAssay(seurat)
)
add_df <- data.frame(Embeddings(seurat, "rca_umap"))
names(add_df) <- paste0("rca_umap", seq(ncol(add_df)))
seurat$rca_umap1 <- add_df$rca_umap1
seurat$rca_umap2 <- add_df$rca_umap2

red_dim_plot(seurat, "rca_umap1", "rca_umap2", "rca_clusters", "cat")
red_dim_plot(seurat, "rca_umap1", "rca_umap2", "cluster_names", "cat")
red_dim_plot(seurat, "rca_umap1", "rca_umap2", "seurat_clusters", "cat")
red_dim_plot(seurat, "rca_umap1", "rca_umap2", "sample", "cat")

cat(
  length(unique(unlist(rca$cell.Type.Estimate))), "cell-types found:",
  unique(unlist(rca$cell.Type.Estimate))
)
cat("Out of", length(colnames(ref)), "possible:", colnames(ref))

for (i in seq_along(markers_trimmed)) {
  if (any(markers_trimmed[[i]] %in% rownames(seurat))) {
    print(names(markers_trimmed[i]))
    print(
      FeaturePlot(
        seurat, markers_trimmed[[i]], reduction = "rca_umap",
        order = TRUE, cols = c("lightgrey", "red")
      )
    )
  }
}
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
