---
title: "03 Subclusters - Combined Organelles"
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "Diana Acosta [Diana.Acosta@osumc.edu]"
  - name: "Hongjun Fu [Hongjun.Fu@osumc.edu]"
  - name: "Ohio State University Wexner Medical Center"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[8], "03_subclusters.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Acosta Organoid scRNAseq Project](https://github.com/eturkes/acosta-organoid-scRNAseq).*

In this document we subcluster annotated cell-types
The data here will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[8]``.

```{r}
#    This file is part of acosta-organoid-scRNAseq.
#    Copyright (C) 2022  Emir Turkes, Diana Acosta, Hongjun Fu, Ohio
#    State University Wexner Medical Center
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

packages <- c(
  "conflicted", "Seurat", "ggplot2", "dplyr", "ggrepel", "SingleCellExperiment", "scuttle", "RCAv2",
  "data.table", "GSVA", "edgeR", "ComplexHeatmap"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
options(future.globals.maxSize = 10485760000)
source(file.path("..", "utils.R"))

protocol <- c("human", "droplet", "single-cell", "umis") # See `cluster_pipeline` in `utils.R`.
vars_to_regress <- NULL # See `cluster_pipeline` in `utils.R`.
parallel_override <- 0 # See `parallel_plan` in `utils.R`.

# Metadata to plot after dimensionality reduction and clustering.
# Values in list can include "no_legend and/or "no_label" to exclude those.
# -------------------------------------------------------------------------
metadata_to_plot <- vector("list", 3)
names(metadata_to_plot) <- c("seurat_clusters", "sample", "timepoint")
# -------------------------------------------------------------------------

analysis_no <- 3
data_name <- unlist(strsplit(getwd(), "/"))[8] # Name of dataset.
data_dir <- file.path("..", "..", "data") # Backed up data.
results_dir <- file.path("..", "..", "results")

# Unique cache directory for each analysis number.
# ------------------------------------------------
cache_dir <- file.path("..", "..", "cache", data_name, paste0("0", analysis_no))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# ------------------------------------------------

knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
```

# Prep

```{r}
ref <- readRDS(file.path(cache_dir, "..", "02", "bhaduri_primary_panel_Subtype.rds"))

seurat <- readRDS(file.path(cache_dir, "..", "02", "rca_seurat.rds"))
seurat

table(seurat$rca_clusters)
remove_clusters <- which(table(seurat$rca_clusters) < 100)
remove <- which(seurat$rca_clusters %in% names(remove_clusters))
cat("Removing", length(remove), "cells across", names(remove_clusters), "cell-types.")

seurat <- seurat[ , -remove]
seurat

red_dim_plot(seurat, "rca_umap1", "rca_umap2", "rca_clusters", "cat")
```

# Excitatory Neuron

## Standard Pipeline

```{r}
sub_seurat <- seurat[ , which(seurat$rca_clusters == "Excitatory Neuron ")]

DefaultAssay(sub_seurat) <- "RNA"
sub_seurat[["SCT"]] <- NULL
sub_name <- "Ex"
sub_seurat <- cluster_pipeline(
  sub_seurat, cache_dir, sub_name, protocol, vars_to_regress, parallel_override, FALSE
)
sub_seurat

for (i in 1:length(metadata_to_plot)) {
  print(names(metadata_to_plot)[i])
  if ("no_legend" %in% metadata_to_plot[[i]] && "no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i]) + NoLegend())
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat") + NoLegend()
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i]))
  } else {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat"))
  }
}
```

## Integration

```{r}
rds <- file.path(cache_dir, paste(sub_name, "integrated_seurat.rds", sep = "_"))
if (file.exists(rds)) {
  integrated_seurat <- readRDS(rds)
} else {

  DefaultAssay(sub_seurat) <- "RNA"
  sub_seurat[["SCT"]] <- NULL

  integrated_seurat <- SplitObject(sub_seurat, "sample")
  integrated_seurat <- lapply(integrated_seurat, SCTransform, verbose = FALSE)
  features <- SelectIntegrationFeatures(integrated_seurat, 3000, verbose = FALSE)
  integrated_seurat <- PrepSCTIntegration(
    integrated_seurat, anchor.features = features, verbose = FALSE
  )
  anchors <- FindIntegrationAnchors(
    integrated_seurat, normalization.method = "SCT", anchor.features = features, verbose = FALSE
  )
  integrated_seurat <- IntegrateData(anchors, normalization.method = "SCT", verbose = FALSE)

  integrated_seurat <- RunPCA(integrated_seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(integrated_seurat)[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  integrated_seurat$pca1 <- add_df$pca1
  integrated_seurat$pca2 <- add_df$pca2

  integrated_seurat <- RunUMAP(integrated_seurat, 1:30, verbose = FALSE)
  add_df <- data.frame(Embeddings(integrated_seurat, "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  integrated_seurat$umap1 <- add_df$umap1
  integrated_seurat$umap2 <- add_df$umap2
  
  integrated_seurat <- FindNeighbors(integrated_seurat, dims = 1:30, verbose = FALSE)
  integrated_seurat <- FindClusters(integrated_seurat, verbose = FALSE)

  saveRDS(integrated_seurat, rds, compress = FALSE)
}

integrated_seurat

for (i in 1:length(metadata_to_plot)) {
  print(names(metadata_to_plot)[i])
  if ("no_legend" %in% metadata_to_plot[[i]] && "no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(integrated_seurat, "umap1", "umap2", names(metadata_to_plot)[i]) + NoLegend()
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        integrated_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat"
      ) + NoLegend()
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(integrated_seurat, "umap1", "umap2", names(metadata_to_plot)[i]))
  } else {
    print(red_dim_plot(integrated_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat"))
  }
}
```

## RCA

```{r}
rds <- file.path(cache_dir, paste(sub_name, "rca.rds", sep = "_"))
if (file.exists(rds)) {
  rca <- readRDS(rds)
} else {
  rca <- createRCAObject(GetAssayData(sub_seurat, "counts"), GetAssayData(sub_seurat))
  rca <- dataProject(
    rca, "Custom", file.path(cache_dir, "..", "02", "bhaduri_primary_panel_Subtype.rds"), "spearman"
  )
  rca <- computeUMAP(rca)
  rca <- estimateCellTypeFromProjection(rca)
  saveRDS(rca, rds, compress = FALSE)
}

rca

sub_seurat$rca_clusters <- factor(unlist(rca$cell.Type.Estimate))
sub_seurat[["rca_umap"]] <- CreateDimReducObject(
  as.matrix(rca$umap.coordinates), key = "rca_umap_", assay = DefaultAssay(sub_seurat)
)
add_df <- data.frame(Embeddings(sub_seurat, "rca_umap"))
names(add_df) <- paste0("rca_umap", seq(ncol(add_df)))
sub_seurat$rca_umap1 <- add_df$rca_umap1
sub_seurat$rca_umap2 <- add_df$rca_umap2

red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "rca_clusters", "cat")
red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "sample", "cat")
red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "timepoint", "cat")

cat(
  length(unique(unlist(rca$cell.Type.Estimate))), "cell-types found:",
  unique(unlist(rca$cell.Type.Estimate))
)
cat("Out of", length(colnames(ref)), "possible:", colnames(ref))
```

## RCA Subset

```{r}
rds <- file.path(cache_dir, paste(sub_name, "rca_sub.rds", sep = "_"))
rds2 <- file.path(cache_dir, paste(sub_name, "panel_sub.rds", sep = "_"))
if (file.exists(rds)) {
  rca <- readRDS(rds)
  ref <- readRDS(rds2)
} else {

  rds <- file.path(cache_dir, "..", "02", "bhaduri_primary_ref.rds")
  if (file.exists(rds)) {
    ref <- readRDS(rds)
  } else {
    ref <- fread(
      file.path(data_dir, "bhaduri_2020", "primary", "exprMatrix.tsv"), data.table = FALSE
    )
    gc()
    rownames(ref) <- ref[ , 1]
    ref <- ref[ , -1]
  
    meta <- read.delim(
      file.path(data_dir, "bhaduri_2020", "primary", "meta.tsv"), row.names = 1
    )
    meta <- meta[-(which(meta$Type == "Outlier")), ]
    meta <- meta[-(which(meta$Subtype == "Microglia low quality")), ]
    meta <- meta[-(which(meta$Subtype == "Low Quality")), ]
  
    ref <- ref[ , which(colnames(ref) %in% rownames(meta))]
  
    ref <- SingleCellExperiment(list(normcounts = ref), colData = meta)
    counts(ref) <- 2 ^ normcounts(ref) - 1
    rm(meta)
    gc()
    saveRDS(ref, rds)
  }
  
  ref <- ref[ , which(ref$Type == "Excitatory Neuron ")]

  rds <- file.path(cache_dir, paste(sub_name, "pseudobulk_sub.rds", sep = "_"))
  if (file.exists(rds)) {
    ref <- readRDS(rds)
  } else {
    top <- nrow(ref) / 2
    top <- ref[1:top, ]
    top <- aggregateAcrossCells(top, top$Subtype, "mean")
  
    bottom <- (nrow(ref) / 2) + 1
    bottom <- ref[bottom:(nrow(ref)), ]
    bottom <- aggregateAcrossCells(bottom, bottom$Subtype, "mean")
  
    ref <- rbind(top, bottom)
    rm(top, bottom)
    saveRDS(ref, rds)
  }

  rds <- file.path(cache_dir, paste(sub_name, "panel_sub.rds", sep = "_"))
  if (file.exists(rds)) {
    ref <- readRDS(rds)
  } else {
    ref <- ref[-(which(rowSums(counts(ref)) == 0)), ]
  
    ref2 <- ref
    colnames(ref2) <- paste(colnames(ref2), "2", sep = "_")
    colnames(ref) <- paste(colnames(ref), "1", sep = "_")
    ref <- cbind(ref, ref2)
    rm(ref2)
  
    ref <- buildReferencePanel(counts(ref))
    saveRDS(ref, rds)
  }

  rds <- file.path(cache_dir, paste(sub_name, "rca_sub.rds", sep = "_"))
  if (file.exists(rds)) {
    rca <- readRDS(rds)
  } else {
    rca <- createRCAObject(GetAssayData(sub_seurat, "counts"), GetAssayData(sub_seurat))
    rca <- dataProject(
      rca, "Custom", file.path(cache_dir, paste(sub_name, "panel_sub.rds", sep = "_")), "spearman"
    )
    rca <- computeUMAP(rca)
    rca <- estimateCellTypeFromProjection(rca)
    saveRDS(rca, rds, compress = FALSE)
  }
}

rca

sub_seurat$rca_clusters <- factor(unlist(rca$cell.Type.Estimate))
sub_seurat[["rca_umap"]] <- CreateDimReducObject(
  as.matrix(rca$umap.coordinates), key = "rca_umap_", assay = DefaultAssay(sub_seurat)
)
add_df <- data.frame(Embeddings(sub_seurat, "rca_umap"))
names(add_df) <- paste0("rca_umap", seq(ncol(add_df)))
sub_seurat$rca_umap1 <- add_df$rca_umap1
sub_seurat$rca_umap2 <- add_df$rca_umap2

red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "rca_clusters", "cat")
red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "sample", "cat")
red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "timepoint", "cat")

cat(
  length(unique(unlist(rca$cell.Type.Estimate))), "cell-types found:",
  unique(unlist(rca$cell.Type.Estimate))
)
cat("Out of", length(colnames(ref)), "possible:", colnames(ref))
```

# Inhibitory Neuron

## Standard Pipeline

```{r}
sub_seurat <- seurat[ , which(seurat$rca_clusters == "Inhibitory Neuron")]

DefaultAssay(sub_seurat) <- "RNA"
sub_seurat[["SCT"]] <- NULL
sub_name <- "In"
sub_seurat <- cluster_pipeline(
  sub_seurat, cache_dir, sub_name, protocol, vars_to_regress, parallel_override, FALSE
)
sub_seurat

for (i in 1:length(metadata_to_plot)) {
  print(names(metadata_to_plot)[i])
  if ("no_legend" %in% metadata_to_plot[[i]] && "no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i]) + NoLegend())
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat") + NoLegend()
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i]))
  } else {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat"))
  }
}
```

## RCA

```{r}
rds <- file.path(cache_dir, paste(sub_name, "rca.rds", sep = "_"))
if (file.exists(rds)) {
  rca <- readRDS(rds)
} else {
  rca <- createRCAObject(GetAssayData(sub_seurat, "counts"), GetAssayData(sub_seurat))
  rca <- dataProject(
    rca, "Custom", file.path(cache_dir, "..", "02", "bhaduri_primary_panel_Subtype.rds"), "spearman"
  )
  rca <- computeUMAP(rca)
  rca <- estimateCellTypeFromProjection(rca)
  saveRDS(rca, rds, compress = FALSE)
}

rca

sub_seurat$rca_clusters <- factor(unlist(rca$cell.Type.Estimate))
sub_seurat[["rca_umap"]] <- CreateDimReducObject(
  as.matrix(rca$umap.coordinates), key = "rca_umap_", assay = DefaultAssay(sub_seurat)
)
add_df <- data.frame(Embeddings(sub_seurat, "rca_umap"))
names(add_df) <- paste0("rca_umap", seq(ncol(add_df)))
sub_seurat$rca_umap1 <- add_df$rca_umap1
sub_seurat$rca_umap2 <- add_df$rca_umap2

red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "rca_clusters", "cat")
red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "sample", "cat")
red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "timepoint", "cat")

cat(
  length(unique(unlist(rca$cell.Type.Estimate))), "cell-types found:",
  unique(unlist(rca$cell.Type.Estimate))
)
cat("Out of", length(colnames(ref)), "possible:", colnames(ref))
```

# IPC

## Standard Pipeline

```{r}
sub_seurat <- seurat[ , which(seurat$rca_clusters == "IPC")]

DefaultAssay(sub_seurat) <- "RNA"
sub_seurat[["SCT"]] <- NULL
sub_name <- "IPC"
sub_seurat <- cluster_pipeline(
  sub_seurat, cache_dir, sub_name, protocol, vars_to_regress, parallel_override, FALSE
)
sub_seurat

for (i in 1:length(metadata_to_plot)) {
  print(names(metadata_to_plot)[i])
  if ("no_legend" %in% metadata_to_plot[[i]] && "no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i]) + NoLegend())
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat") + NoLegend()
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i]))
  } else {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat"))
  }
}
```

## Integration

```{r}
rds <- file.path(cache_dir, paste(sub_name, "integrated_seurat.rds", sep = "_"))
if (file.exists(rds)) {
  integrated_seurat <- readRDS(rds)
} else {

  DefaultAssay(sub_seurat) <- "RNA"
  sub_seurat[["SCT"]] <- NULL

  integrated_seurat <- SplitObject(sub_seurat, "sample")
  integrated_seurat <- lapply(integrated_seurat, SCTransform, verbose = FALSE)
  features <- SelectIntegrationFeatures(integrated_seurat, 3000, verbose = FALSE)
  integrated_seurat <- PrepSCTIntegration(
    integrated_seurat, anchor.features = features, verbose = FALSE
  )
  anchors <- FindIntegrationAnchors(
    integrated_seurat, normalization.method = "SCT", anchor.features = features, verbose = FALSE
  )
  integrated_seurat <- IntegrateData(anchors, normalization.method = "SCT", verbose = FALSE)

  integrated_seurat <- RunPCA(integrated_seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(integrated_seurat)[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  integrated_seurat$pca1 <- add_df$pca1
  integrated_seurat$pca2 <- add_df$pca2

  integrated_seurat <- RunUMAP(integrated_seurat, 1:30, verbose = FALSE)
  add_df <- data.frame(Embeddings(integrated_seurat, "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  integrated_seurat$umap1 <- add_df$umap1
  integrated_seurat$umap2 <- add_df$umap2
  
  integrated_seurat <- FindNeighbors(integrated_seurat, dims = 1:30, verbose = FALSE)
  integrated_seurat <- FindClusters(integrated_seurat, verbose = FALSE)

  saveRDS(integrated_seurat, rds, compress = FALSE)
}

integrated_seurat

for (i in 1:length(metadata_to_plot)) {
  print(names(metadata_to_plot)[i])
  if ("no_legend" %in% metadata_to_plot[[i]] && "no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(integrated_seurat, "umap1", "umap2", names(metadata_to_plot)[i]) + NoLegend()
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        integrated_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat"
      ) + NoLegend()
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(integrated_seurat, "umap1", "umap2", names(metadata_to_plot)[i]))
  } else {
    print(red_dim_plot(integrated_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat"))
  }
}
```

## RCA

```{r}
rds <- file.path(cache_dir, paste(sub_name, "rca.rds", sep = "_"))
if (file.exists(rds)) {
  rca <- readRDS(rds)
} else {
  rca <- createRCAObject(GetAssayData(sub_seurat, "counts"), GetAssayData(sub_seurat))
  rca <- dataProject(
    rca, "Custom", file.path(cache_dir, "..", "02", "bhaduri_primary_panel_Subtype.rds"), "spearman"
  )
  rca <- computeUMAP(rca)
  rca <- estimateCellTypeFromProjection(rca)
  saveRDS(rca, rds, compress = FALSE)
}

rca

sub_seurat$rca_clusters <- factor(unlist(rca$cell.Type.Estimate))
sub_seurat[["rca_umap"]] <- CreateDimReducObject(
  as.matrix(rca$umap.coordinates), key = "rca_umap_", assay = DefaultAssay(sub_seurat)
)
add_df <- data.frame(Embeddings(sub_seurat, "rca_umap"))
names(add_df) <- paste0("rca_umap", seq(ncol(add_df)))
sub_seurat$rca_umap1 <- add_df$rca_umap1
sub_seurat$rca_umap2 <- add_df$rca_umap2

red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "rca_clusters", "cat")
red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "sample", "cat")
red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "timepoint", "cat")

cat(
  length(unique(unlist(rca$cell.Type.Estimate))), "cell-types found:",
  unique(unlist(rca$cell.Type.Estimate))
)
cat("Out of", length(colnames(ref)), "possible:", colnames(ref))
```

## RCA Subset

```{r}
rds <- file.path(cache_dir, paste(sub_name, "rca_sub.rds", sep = "_"))
rds2 <- file.path(cache_dir, paste(sub_name, "panel_sub.rds", sep = "_"))
if (file.exists(rds)) {
  rca <- readRDS(rds)
  ref <- readRDS(rds2)
} else {

  rds <- file.path(cache_dir, "..", "02", "bhaduri_primary_ref.rds")
  if (file.exists(rds)) {
    ref <- readRDS(rds)
  } else {
    ref <- fread(
      file.path(data_dir, "bhaduri_2020", "primary", "exprMatrix.tsv"), data.table = FALSE
    )
    gc()
    rownames(ref) <- ref[ , 1]
    ref <- ref[ , -1]
  
    meta <- read.delim(
      file.path(data_dir, "bhaduri_2020", "primary", "meta.tsv"), row.names = 1
    )
    meta <- meta[-(which(meta$Type == "Outlier")), ]
    meta <- meta[-(which(meta$Subtype == "Microglia low quality")), ]
    meta <- meta[-(which(meta$Subtype == "Low Quality")), ]
  
    ref <- ref[ , which(colnames(ref) %in% rownames(meta))]
  
    ref <- SingleCellExperiment(list(normcounts = ref), colData = meta)
    counts(ref) <- 2 ^ normcounts(ref) - 1
    rm(meta)
    gc()
    saveRDS(ref, rds)
  }
  
  ref <- ref[ , which(ref$Type == "IPC")]

  rds <- file.path(cache_dir, paste(sub_name, "pseudobulk_sub.rds", sep = "_"))
  if (file.exists(rds)) {
    ref <- readRDS(rds)
  } else {
    top <- nrow(ref) / 2
    top <- ref[1:top, ]
    top <- aggregateAcrossCells(top, top$Subtype, "mean")
  
    bottom <- (nrow(ref) / 2) + 1
    bottom <- ref[bottom:(nrow(ref)), ]
    bottom <- aggregateAcrossCells(bottom, bottom$Subtype, "mean")
  
    ref <- rbind(top, bottom)
    rm(top, bottom)
    saveRDS(ref, rds)
  }

  rds <- file.path(cache_dir, paste(sub_name, "panel_sub.rds", sep = "_"))
  if (file.exists(rds)) {
    ref <- readRDS(rds)
  } else {
    ref <- ref[-(which(rowSums(counts(ref)) == 0)), ]
  
    ref2 <- ref
    colnames(ref2) <- paste(colnames(ref2), "2", sep = "_")
    colnames(ref) <- paste(colnames(ref), "1", sep = "_")
    ref <- cbind(ref, ref2)
    rm(ref2)
  
    ref <- buildReferencePanel(counts(ref))
    saveRDS(ref, rds)
  }

  rds <- file.path(cache_dir, paste(sub_name, "rca_sub.rds", sep = "_"))
  if (file.exists(rds)) {
    rca <- readRDS(rds)
  } else {
    rca <- createRCAObject(GetAssayData(sub_seurat, "counts"), GetAssayData(sub_seurat))
    rca <- dataProject(
      rca, "Custom", file.path(cache_dir, paste(sub_name, "panel_sub.rds", sep = "_")), "spearman"
    )
    rca <- computeUMAP(rca)
    rca <- estimateCellTypeFromProjection(rca)
    saveRDS(rca, rds, compress = FALSE)
  }
}

rca

sub_seurat$rca_clusters <- factor(unlist(rca$cell.Type.Estimate))
sub_seurat[["rca_umap"]] <- CreateDimReducObject(
  as.matrix(rca$umap.coordinates), key = "rca_umap_", assay = DefaultAssay(sub_seurat)
)
add_df <- data.frame(Embeddings(sub_seurat, "rca_umap"))
names(add_df) <- paste0("rca_umap", seq(ncol(add_df)))
sub_seurat$rca_umap1 <- add_df$rca_umap1
sub_seurat$rca_umap2 <- add_df$rca_umap2

red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "rca_clusters", "cat")
red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "sample", "cat")
red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "timepoint", "cat")

cat(
  length(unique(unlist(rca$cell.Type.Estimate))), "cell-types found:",
  unique(unlist(rca$cell.Type.Estimate))
)
cat("Out of", length(colnames(ref)), "possible:", colnames(ref))
```

# Mural

## Standard Pipeline

```{r}
sub_seurat <- seurat[ , which(seurat$rca_clusters == "Mural")]

DefaultAssay(sub_seurat) <- "RNA"
sub_seurat[["SCT"]] <- NULL
sub_name <- "Mu"
sub_seurat <- cluster_pipeline(
  sub_seurat, cache_dir, sub_name, protocol, vars_to_regress, parallel_override, FALSE
)
sub_seurat

for (i in 1:length(metadata_to_plot)) {
  print(names(metadata_to_plot)[i])
  if ("no_legend" %in% metadata_to_plot[[i]] && "no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i]) + NoLegend())
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat") + NoLegend()
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i]))
  } else {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat"))
  }
}
```

## RCA

```{r}
rds <- file.path(cache_dir, paste(sub_name, "rca.rds", sep = "_"))
if (file.exists(rds)) {
  rca <- readRDS(rds)
} else {
  rca <- createRCAObject(GetAssayData(sub_seurat, "counts"), GetAssayData(sub_seurat))
  rca <- dataProject(
    rca, "Custom", file.path(cache_dir, "..", "02", "bhaduri_primary_panel_Subtype.rds"), "spearman"
  )
  rca <- computeUMAP(rca)
  rca <- estimateCellTypeFromProjection(rca)
  saveRDS(rca, rds, compress = FALSE)
}

rca

sub_seurat$rca_clusters <- factor(unlist(rca$cell.Type.Estimate))
sub_seurat[["rca_umap"]] <- CreateDimReducObject(
  as.matrix(rca$umap.coordinates), key = "rca_umap_", assay = DefaultAssay(sub_seurat)
)
add_df <- data.frame(Embeddings(sub_seurat, "rca_umap"))
names(add_df) <- paste0("rca_umap", seq(ncol(add_df)))
sub_seurat$rca_umap1 <- add_df$rca_umap1
sub_seurat$rca_umap2 <- add_df$rca_umap2

red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "rca_clusters", "cat")
red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "sample", "cat")
red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "timepoint", "cat")

cat(
  length(unique(unlist(rca$cell.Type.Estimate))), "cell-types found:",
  unique(unlist(rca$cell.Type.Estimate))
)
cat("Out of", length(colnames(ref)), "possible:", colnames(ref))
```

# Radial Glia

## Standard Pipeline

```{r}
sub_seurat <- seurat[ , which(seurat$rca_clusters == "Radial Glia")]

DefaultAssay(sub_seurat) <- "RNA"
sub_seurat[["SCT"]] <- NULL
sub_name <- "RG"
sub_seurat <- cluster_pipeline(
  sub_seurat, cache_dir, sub_name, protocol, vars_to_regress, parallel_override, FALSE
)
sub_seurat

for (i in 1:length(metadata_to_plot)) {
  print(names(metadata_to_plot)[i])
  if ("no_legend" %in% metadata_to_plot[[i]] && "no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i]) + NoLegend())
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat") + NoLegend()
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i]))
  } else {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat"))
  }
}
```

## Integration

```{r}
rds <- file.path(cache_dir, paste(sub_name, "integrated_seurat.rds", sep = "_"))
if (file.exists(rds)) {
  integrated_seurat <- readRDS(rds)
} else {

  DefaultAssay(sub_seurat) <- "RNA"
  sub_seurat[["SCT"]] <- NULL

  integrated_seurat <- SplitObject(sub_seurat, "sample")
  integrated_seurat <- lapply(integrated_seurat, SCTransform, verbose = FALSE)
  features <- SelectIntegrationFeatures(integrated_seurat, 3000, verbose = FALSE)
  integrated_seurat <- PrepSCTIntegration(
    integrated_seurat, anchor.features = features, verbose = FALSE
  )
  anchors <- FindIntegrationAnchors(
    integrated_seurat, normalization.method = "SCT", anchor.features = features, verbose = FALSE
  )
  integrated_seurat <- IntegrateData(anchors, normalization.method = "SCT", verbose = FALSE)

  integrated_seurat <- RunPCA(integrated_seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(integrated_seurat)[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  integrated_seurat$pca1 <- add_df$pca1
  integrated_seurat$pca2 <- add_df$pca2

  integrated_seurat <- RunUMAP(integrated_seurat, 1:30, verbose = FALSE)
  add_df <- data.frame(Embeddings(integrated_seurat, "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  integrated_seurat$umap1 <- add_df$umap1
  integrated_seurat$umap2 <- add_df$umap2

  integrated_seurat <- FindNeighbors(integrated_seurat, dims = 1:30, verbose = FALSE)
  integrated_seurat <- FindClusters(integrated_seurat, verbose = FALSE)

  saveRDS(integrated_seurat, rds, compress = FALSE)
}

integrated_seurat

for (i in 1:length(metadata_to_plot)) {
  print(names(metadata_to_plot)[i])
  if ("no_legend" %in% metadata_to_plot[[i]] && "no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(integrated_seurat, "umap1", "umap2", names(metadata_to_plot)[i]) + NoLegend()
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        integrated_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat"
      ) + NoLegend()
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(integrated_seurat, "umap1", "umap2", names(metadata_to_plot)[i]))
  } else {
    print(red_dim_plot(integrated_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat"))
  }
}
```

## RCA

```{r}
rds <- file.path(cache_dir, paste(sub_name, "rca.rds", sep = "_"))
if (file.exists(rds)) {
  rca <- readRDS(rds)
} else {
  rca <- createRCAObject(GetAssayData(sub_seurat, "counts"), GetAssayData(sub_seurat))
  rca <- dataProject(
    rca, "Custom", file.path(cache_dir, "..", "02", "bhaduri_primary_panel_Subtype.rds"), "spearman"
  )
  rca <- computeUMAP(rca)
  rca <- estimateCellTypeFromProjection(rca)
  saveRDS(rca, rds, compress = FALSE)
}

rca

sub_seurat$rca_clusters <- factor(unlist(rca$cell.Type.Estimate))
sub_seurat[["rca_umap"]] <- CreateDimReducObject(
  as.matrix(rca$umap.coordinates), key = "rca_umap_", assay = DefaultAssay(sub_seurat)
)
add_df <- data.frame(Embeddings(sub_seurat, "rca_umap"))
names(add_df) <- paste0("rca_umap", seq(ncol(add_df)))
sub_seurat$rca_umap1 <- add_df$rca_umap1
sub_seurat$rca_umap2 <- add_df$rca_umap2

red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "rca_clusters", "cat")
red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "sample", "cat")
red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "timepoint", "cat")

cat(
  length(unique(unlist(rca$cell.Type.Estimate))), "cell-types found:",
  unique(unlist(rca$cell.Type.Estimate))
)
cat("Out of", length(colnames(ref)), "possible:", colnames(ref))
```

## RCA Subset

```{r}
rds <- file.path(cache_dir, paste(sub_name, "rca_sub.rds", sep = "_"))
rds2 <- file.path(cache_dir, paste(sub_name, "panel_sub.rds", sep = "_"))
if (file.exists(rds)) {
  rca <- readRDS(rds)
  ref <- readRDS(rds2)
} else {

  rds <- file.path(cache_dir, "..", "02", "bhaduri_primary_ref.rds")
  if (file.exists(rds)) {
    ref <- readRDS(rds)
  } else {
    ref <- fread(
      file.path(data_dir, "bhaduri_2020", "primary", "exprMatrix.tsv"), data.table = FALSE
    )
    gc()
    rownames(ref) <- ref[ , 1]
    ref <- ref[ , -1]
  
    meta <- read.delim(
      file.path(data_dir, "bhaduri_2020", "primary", "meta.tsv"), row.names = 1
    )
    meta <- meta[-(which(meta$Type == "Outlier")), ]
    meta <- meta[-(which(meta$Subtype == "Microglia low quality")), ]
    meta <- meta[-(which(meta$Subtype == "Low Quality")), ]
  
    ref <- ref[ , which(colnames(ref) %in% rownames(meta))]
  
    ref <- SingleCellExperiment(list(normcounts = ref), colData = meta)
    counts(ref) <- 2 ^ normcounts(ref) - 1
    rm(meta)
    gc()
    saveRDS(ref, rds)
  }
  
  ref <- ref[ , which(ref$Type == "Radial Glia")]

  rds <- file.path(cache_dir, paste(sub_name, "pseudobulk_sub.rds", sep = "_"))
  if (file.exists(rds)) {
    ref <- readRDS(rds)
  } else {
    top <- nrow(ref) / 2
    top <- ref[1:top, ]
    top <- aggregateAcrossCells(top, top$Subtype, "mean")
  
    bottom <- (nrow(ref) / 2) + 1
    bottom <- ref[bottom:(nrow(ref)), ]
    bottom <- aggregateAcrossCells(bottom, bottom$Subtype, "mean")
  
    ref <- rbind(top, bottom)
    rm(top, bottom)
    saveRDS(ref, rds)
  }

  rds <- file.path(cache_dir, paste(sub_name, "panel_sub.rds", sep = "_"))
  if (file.exists(rds)) {
    ref <- readRDS(rds)
  } else {
    ref <- ref[-(which(rowSums(counts(ref)) == 0)), ]
  
    ref2 <- ref
    colnames(ref2) <- paste(colnames(ref2), "2", sep = "_")
    colnames(ref) <- paste(colnames(ref), "1", sep = "_")
    ref <- cbind(ref, ref2)
    rm(ref2)
  
    ref <- buildReferencePanel(counts(ref))
    saveRDS(ref, rds)
  }

  rds <- file.path(cache_dir, paste(sub_name, "rca_sub.rds", sep = "_"))
  if (file.exists(rds)) {
    rca <- readRDS(rds)
  } else {
    rca <- createRCAObject(GetAssayData(sub_seurat, "counts"), GetAssayData(sub_seurat))
    rca <- dataProject(
      rca, "Custom", file.path(cache_dir, paste(sub_name, "panel_sub.rds", sep = "_")), "spearman"
    )
    rca <- computeUMAP(rca)
    rca <- estimateCellTypeFromProjection(rca)
    saveRDS(rca, rds, compress = FALSE)
  }
}

rca

sub_seurat$rca_clusters <- factor(unlist(rca$cell.Type.Estimate))
sub_seurat[["rca_umap"]] <- CreateDimReducObject(
  as.matrix(rca$umap.coordinates), key = "rca_umap_", assay = DefaultAssay(sub_seurat)
)
add_df <- data.frame(Embeddings(sub_seurat, "rca_umap"))
names(add_df) <- paste0("rca_umap", seq(ncol(add_df)))
sub_seurat$rca_umap1 <- add_df$rca_umap1
sub_seurat$rca_umap2 <- add_df$rca_umap2

red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "rca_clusters", "cat")
red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "sample", "cat")
red_dim_plot(sub_seurat, "rca_umap1", "rca_umap2", "timepoint", "cat")

cat(
  length(unique(unlist(rca$cell.Type.Estimate))), "cell-types found:",
  unique(unlist(rca$cell.Type.Estimate))
)
cat("Out of", length(colnames(ref)), "possible:", colnames(ref))
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
