---
title: "04 DEGs - Combined Organelles"
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "Diana Acosta [Diana.Acosta@osumc.edu]"
  - name: "Hongjun Fu [Hongjun.Fu@osumc.edu]"
  - name: "Ohio State University Wexner Medical Center"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[8], "04_DEGs.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
.tocify-subheader .tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 45px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Acosta Organoid scRNAseq Project](https://github.com/eturkes/acosta-organoid-scRNAseq).*

In this document we subcluster annotated cell-types
The data here will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[8]``.

```{r}
#    This file is part of acosta-organoid-scRNAseq.
#    Copyright (C) 2022  Emir Turkes, Diana Acosta, Hongjun Fu, Ohio
#    State University Wexner Medical Center
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

packages <- c(
  "conflicted", "Seurat", "scuttle", "edgeR", "DT", "ComplexHeatmap", "VennDiagram", "RColorBrewer"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path("..", "utils.R"))
conflict_prefer("cpm", "edgeR", quiet = TRUE)

analysis_no <- 4
data_name <- unlist(strsplit(getwd(), "/"))[8] # Name of dataset.
data_dir <- file.path("..", "..", "data") # Backed up data.
results_dir <- file.path("..", "..", "results")

# Unique cache directory for each analysis number.
# ------------------------------------------------
cache_dir <- file.path("..", "..", "cache", data_name, "03")
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# ------------------------------------------------

knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
```

# Excitatory Neuron

```{r}
sub_name <- "Ex"
seurat_list <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, "seurat_list.rds", sep = "_"))
)
```

## D120

```{r}
seurat_idx <- 1

seurat_list[[seurat_idx]]@active.ident <- factor(
  seurat_list[[seurat_idx]]$genotype, c("WT", "HET", "HOM")
)
```

### All

```{r}
sce <- SingleCellExperiment(list(counts = GetAssayData(seurat_list[[seurat_idx]], slot = "counts")))
sce$sample <- seurat_list[[seurat_idx]]$sample
sce$genotype <- seurat_list[[seurat_idx]]$genotype
sce$clustermap_names <- seurat_list[[seurat_idx]]$clustermap_names

sce <- suppressWarnings(aggregateAcrossCells(sce, sce$sample))
keep <- filterByExpr(counts(sce), group = sce$genotype, min.total.count = 10)
sce <- sce[keep, ]
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

design <- model.matrix(~ 0 + sce$genotype)
colnames(design) <- unique(sce$genotype)
fit <- lmFit(logcounts(sce), design)
contrast_mat <- makeContrasts(HOM-HET, HET-WT, HOM-WT, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)
tests <- decideTests(cont_fit, "nestedF", p.value = 0.05)
```

#### HOM-HET, HET-WT, HOM-WT

```{r}
tests_up <- tests[tests[ , 1] == 1 & tests[ , 2] == 1 & tests[ , 3] == 1, ]
tests_down <- tests[tests[ , 1] == -1 & tests[ , 2] == -1 & tests[ , 3] == -1, ]

results <- topTable(cont_fit, number = Inf)
up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$F), ]
}

datatable_download_exp(all)

mat <- logcounts(sce)[which(rownames(sce) %in% rownames(all)[1:25]), ]
mat <- t(apply(mat, 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    column_order = c(5, 6, 1, 2, 3, 4),
    column_names_rot = 45,
    heatmap_legend_param = list(
      title = "Per-row Scaled Pseudobulk Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)

DotPlot(
  seurat_list[[seurat_idx]], features = rownames(all)[1:25], cols = c("blue", "red")
) + RotatedAxis()
VlnPlot(seurat_list[[seurat_idx]], rownames(all)[1:12])
```

#### HET-WT, HOM-WT

```{r}
tests_up <- tests[tests[ , 2] == 1 & tests[ , 3] == 1, ]
tests_down <- tests[tests[ , 2] == -1 & tests[ , 3] == -1, ]

results <- topTable(cont_fit, c(2, 3), number = Inf)
up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$F), ]
}

datatable_download_exp(all)

mat <- logcounts(sce)[which(rownames(sce) %in% rownames(all)[1:25]), ]
mat <- t(apply(mat, 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    column_order = c(5, 6, 1, 2, 3, 4),
    column_names_rot = 45,
    heatmap_legend_param = list(
      title = "Per-row Scaled Pseudobulk Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)

DotPlot(
  seurat_list[[seurat_idx]], features = rownames(all)[1:25], cols = c("blue", "red")
) + RotatedAxis()
VlnPlot(seurat_list[[seurat_idx]], rownames(all)[1:12])
```

#### HOM-HET, HOM-WT

```{r}
tests_up <- tests[tests[ , 1] == 1 & tests[ , 3] == 1, ]
tests_down <- tests[tests[ , 1] == -1 & tests[ , 3] == -1, ]

results <- topTable(cont_fit, c(1, 3), number = Inf)
up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$F), ]
}

datatable_download_exp(all)

mat <- logcounts(sce)[which(rownames(sce) %in% rownames(all)[1:25]), ]
mat <- t(apply(mat, 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    column_order = c(5, 6, 1, 2, 3, 4),
    column_names_rot = 45,
    heatmap_legend_param = list(
      title = "Per-row Scaled Pseudobulk Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)

DotPlot(
  seurat_list[[seurat_idx]], features = rownames(all)[1:25], cols = c("blue", "red")
) + RotatedAxis()
VlnPlot(seurat_list[[seurat_idx]], rownames(all)[1:12])
```

#### HOM-HET

```{r}
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]

results <- topTable(cont_fit, 1, number = Inf)
up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}

datatable_download_exp(all)

mat <- logcounts(sce)[which(rownames(sce) %in% rownames(all)[1:25]), ]
mat <- t(apply(mat, 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    column_order = c(5, 6, 1, 2, 3, 4),
    column_names_rot = 45,
    heatmap_legend_param = list(
      title = "Per-row Scaled Pseudobulk Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)

DotPlot(
  seurat_list[[seurat_idx]], features = rownames(all)[1:25], cols = c("blue", "red")
) + RotatedAxis()
VlnPlot(seurat_list[[seurat_idx]], rownames(all)[1:12])
```

#### HET-WT

```{r}
tests_up <- tests[tests[ , 2] == 1, ]
tests_down <- tests[tests[ , 2] == -1, ]

results <- topTable(cont_fit, 2, number = Inf)
up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}

datatable_download_exp(all)

mat <- logcounts(sce)[which(rownames(sce) %in% rownames(all)[1:25]), ]
mat <- t(apply(mat, 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    column_order = c(5, 6, 1, 2, 3, 4),
    column_names_rot = 45,
    heatmap_legend_param = list(
      title = "Per-row Scaled Pseudobulk Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)

DotPlot(
  seurat_list[[seurat_idx]], features = rownames(all)[1:25], cols = c("blue", "red")
) + RotatedAxis()
VlnPlot(seurat_list[[seurat_idx]], rownames(all)[1:12])
```

#### HOM-WT

```{r}
tests_up <- tests[tests[ , 3] == 1, ]
tests_down <- tests[tests[ , 3] == -1, ]

results <- topTable(cont_fit, 3, number = Inf)
up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}

datatable_download_exp(all)

mat <- logcounts(sce)[which(rownames(sce) %in% rownames(all)[1:25]), ]
mat <- t(apply(mat, 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    column_order = c(5, 6, 1, 2, 3, 4),
    column_names_rot = 45,
    heatmap_legend_param = list(
      title = "Per-row Scaled Pseudobulk Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)

DotPlot(
  seurat_list[[seurat_idx]], features = rownames(all)[1:25], cols = c("blue", "red")
) + RotatedAxis()
VlnPlot(seurat_list[[seurat_idx]], rownames(all)[1:12])
```

#### Using `Seurat::FindClusters` for HOM-WT

```{r}
seurat_DEGs <- FindMarkers(seurat_list[[seurat_idx]], "HOM", "WT", verbose = FALSE)
datatable_download_exp(seurat_DEGs)
VlnPlot(seurat_list[[seurat_idx]], rownames(seurat_DEGs)[1:12])

grid.newpage()
plot <- venn.diagram(
  list(rownames(all), rownames(seurat_DEGs)), NULL,
  TRUE, category.names = c("Limma Trend", "Seurat FindClusters"),
  fill = brewer.pal(3, "Pastel2")[1:2]
)
grid.draw(plot)
```

#### Using a simpler limma model for HOM-WT

```{r}
contrast_mat <- makeContrasts(HOM-HET, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)
results <- topTable(cont_fit, number = Inf, p.value = 0.05)

datatable_download_exp(results)

grid.newpage()
plot <- venn.diagram(
  list(rownames(results), rownames(seurat_DEGs)), NULL,
  TRUE, category.names = c("Limma Trend", "Seurat FindClusters"),
  fill = brewer.pal(3, "Pastel2")[1:2]
)
grid.draw(plot)
```

### Ex Upper Layer 1

```{r}
sub_seurat <- subset(seurat_list[[seurat_idx]], clustermap_names == "Ex Upper Layer 1")

sce <- SingleCellExperiment(list(counts = GetAssayData(sub_seurat, slot = "counts")))
sce$sample <- sub_seurat$sample
sce$genotype <- sub_seurat$genotype
sce$clustermap_names <- sub_seurat$clustermap_names

sce <- suppressWarnings(aggregateAcrossCells(sce, sce$sample))
keep <- filterByExpr(counts(sce), group = sce$genotype, min.total.count = 10)
sce <- sce[keep, ]
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

design <- model.matrix(~ 0 + sce$genotype)
colnames(design) <- unique(sce$genotype)
fit <- lmFit(logcounts(sce), design)
contrast_mat <- makeContrasts(HOM-HET, HET-WT, HOM-WT, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)
tests <- decideTests(cont_fit, "nestedF", p.value = 0.05)
```

#### HOM-HET, HET-WT, HOM-WT

```{r}
tests_up <- tests[tests[ , 1] == 1 & tests[ , 2] == 1 & tests[ , 3] == 1, ]
tests_down <- tests[tests[ , 1] == -1 & tests[ , 2] == -1 & tests[ , 3] == -1, ]

results <- topTable(cont_fit, number = Inf)
up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$F), ]
}

datatable_download_exp(all)

mat <- logcounts(sce)[which(rownames(sce) %in% rownames(all)[1:25]), ]
mat <- t(apply(mat, 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    column_order = c(5, 6, 1, 2, 3, 4),
    column_names_rot = 45,
    heatmap_legend_param = list(
      title = "Per-row Scaled Pseudobulk Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)

DotPlot(sub_seurat, features = rownames(all)[1:25], cols = c("blue", "red")) + RotatedAxis()
VlnPlot(sub_seurat, rownames(all)[1:12])
```

#### HET-WT, HOM-WT

```{r}
tests_up <- tests[tests[ , 2] == 1 & tests[ , 3] == 1, ]
tests_down <- tests[tests[ , 2] == -1 & tests[ , 3] == -1, ]

results <- topTable(cont_fit, c(2, 3), number = Inf)
up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$F), ]
}

datatable_download_exp(all)

mat <- logcounts(sce)[which(rownames(sce) %in% rownames(all)[1:25]), ]
mat <- t(apply(mat, 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    column_order = c(5, 6, 1, 2, 3, 4),
    column_names_rot = 45,
    heatmap_legend_param = list(
      title = "Per-row Scaled Pseudobulk Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)

DotPlot(sub_seurat, features = rownames(all)[1:25], cols = c("blue", "red")) + RotatedAxis()
VlnPlot(sub_seurat, rownames(all)[1:12])
```

#### HOM-HET, HOM-WT

```{r}
tests_up <- tests[tests[ , 1] == 1 & tests[ , 3] == 1, ]
tests_down <- tests[tests[ , 1] == -1 & tests[ , 3] == -1, ]

results <- topTable(cont_fit, c(1, 3), number = Inf)
up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$F), ]
}

datatable_download_exp(all)

mat <- logcounts(sce)[which(rownames(sce) %in% rownames(all)[1:25]), ]
mat <- t(apply(mat, 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    column_order = c(5, 6, 1, 2, 3, 4),
    column_names_rot = 45,
    heatmap_legend_param = list(
      title = "Per-row Scaled Pseudobulk Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)

DotPlot(sub_seurat, features = rownames(all)[1:25], cols = c("blue", "red")) + RotatedAxis()
VlnPlot(sub_seurat, rownames(all)[1:12])
```

#### HOM-HET

```{r}
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]

results <- topTable(cont_fit, 1, number = Inf)
up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}

datatable_download_exp(all)

mat <- logcounts(sce)[which(rownames(sce) %in% rownames(all)[1:25]), ]
mat <- t(apply(mat, 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    column_order = c(5, 6, 1, 2, 3, 4),
    column_names_rot = 45,
    heatmap_legend_param = list(
      title = "Per-row Scaled Pseudobulk Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)

DotPlot(sub_seurat, features = rownames(all)[1:25], cols = c("blue", "red")) + RotatedAxis()
VlnPlot(sub_seurat, rownames(all)[1:12])
```

#### HET-WT

```{r}
tests_up <- tests[tests[ , 2] == 1, ]
tests_down <- tests[tests[ , 2] == -1, ]

results <- topTable(cont_fit, 2, number = Inf)
up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}

datatable_download_exp(all)

mat <- logcounts(sce)[which(rownames(sce) %in% rownames(all)[1:25]), ]
mat <- t(apply(mat, 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    column_order = c(5, 6, 1, 2, 3, 4),
    column_names_rot = 45,
    heatmap_legend_param = list(
      title = "Per-row Scaled Pseudobulk Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)

DotPlot(sub_seurat, features = rownames(all)[1:25], cols = c("blue", "red")) + RotatedAxis()
VlnPlot(sub_seurat, rownames(all)[1:12])
```

#### HOM-WT

```{r}
tests_up <- tests[tests[ , 3] == 1, ]
tests_down <- tests[tests[ , 3] == -1, ]

results <- topTable(cont_fit, 3, number = Inf)
up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}

datatable_download_exp(all)

mat <- logcounts(sce)[which(rownames(sce) %in% rownames(all)[1:25]), ]
mat <- t(apply(mat, 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    column_order = c(5, 6, 1, 2, 3, 4),
    column_names_rot = 45,
    heatmap_legend_param = list(
      title = "Per-row Scaled Pseudobulk Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)

DotPlot(sub_seurat, features = rownames(all)[1:25], cols = c("blue", "red")) + RotatedAxis()
VlnPlot(sub_seurat, rownames(all)[1:12])
```

#### Using `Seurat::FindClusters` for HOM-WT

```{r}
seurat_DEGs <- FindMarkers(seurat_list[[seurat_idx]], "HOM", "WT", verbose = FALSE)
datatable_download_exp(seurat_DEGs)
VlnPlot(seurat_list[[seurat_idx]], rownames(seurat_DEGs)[1:12])

grid.newpage()
plot <- venn.diagram(
  list(rownames(all), rownames(seurat_DEGs)), NULL,
  TRUE, category.names = c("Limma Trend", "Seurat FindClusters"),
  fill = brewer.pal(3, "Pastel2")[1:2]
)
grid.draw(plot)
```

#### Using a simpler limma model for HOM-WT

```{r}
contrast_mat <- makeContrasts(HOM-HET, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)
results <- topTable(cont_fit, number = Inf, p.value = 0.05)

datatable_download_exp(results)

grid.newpage()
plot <- venn.diagram(
  list(rownames(results), rownames(seurat_DEGs)), NULL,
  TRUE, category.names = c("Limma Trend", "Seurat FindClusters"),
  fill = brewer.pal(3, "Pastel2")[1:2]
)
grid.draw(plot)
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
