---
title: "04 Subcluster Annotations - Combined Organelles"
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "Diana Acosta [Diana.Acosta@osumc.edu]"
  - name: "Hongjun Fu [Hongjun.Fu@osumc.edu]"
  - name: "Ohio State University Wexner Medical Center"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[8], "04_subcluster_annotations.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Acosta Organoid scRNAseq Project](https://github.com/eturkes/acosta-organoid-scRNAseq).*

In this document we subcluster annotated cell-types
The data here will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[8]``.

```{r}
#    This file is part of acosta-organoid-scRNAseq.
#    Copyright (C) 2022  Emir Turkes, Diana Acosta, Hongjun Fu, Ohio
#    State University Wexner Medical Center
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

packages <- c("conflicted", "Seurat", "ggplot2", "dplyr", "ggrepel", "DT")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path("..", "utils.R"))

analysis_no <- 4
data_name <- unlist(strsplit(getwd(), "/"))[8] # Name of dataset.
data_dir <- file.path("..", "..", "data") # Backed up data.
results_dir <- file.path("..", "..", "results")

# Unique cache directory for each analysis number.
# ------------------------------------------------
cache_dir <- file.path("..", "..", "cache", data_name, paste0("0", analysis_no))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# ------------------------------------------------

knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
```

# Prep

```{r}
timepoints <- vector("list", 2)
names(timepoints) <- c("D72", "D120")
```

# Excitatory Neurons

```{r}
sub_name <- "Ex"
```

## D72

```{r}
timepoints[[1]] <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, names(timepoints)[1], "seurat.rds", sep = "_"))
)

rds <- file.path(cache_dir, paste(sub_name, names(timepoints)[1], "markers.rds", sep = "_"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(timepoints[[1]], only.pos = TRUE, verbose = FALSE)
  rownames(markers) <- NULL
  saveRDS(markers, rds)
}

datatable_download(markers)
```

## D120

```{r}
timepoints[[2]] <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, names(timepoints)[2], "seurat.rds", sep = "_"))
)

rds <- file.path(cache_dir, paste(sub_name, names(timepoints)[2], "markers.rds", sep = "_"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(timepoints[[2]], only.pos = TRUE, verbose = FALSE)
  rownames(markers) <- NULL
  saveRDS(markers, rds)
}

datatable_download(markers)
```

# Inhibitory Neurons

```{r}
sub_name <- "In"
```

## D72

```{r}
timepoints[[1]] <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, names(timepoints)[1], "seurat.rds", sep = "_"))
)

rds <- file.path(cache_dir, paste(sub_name, names(timepoints)[1], "markers.rds", sep = "_"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(timepoints[[1]], only.pos = TRUE, verbose = FALSE)
  rownames(markers) <- NULL
  saveRDS(markers, rds)
}

datatable_download(markers)
```

## D120

```{r}
timepoints[[2]] <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, names(timepoints)[2], "seurat.rds", sep = "_"))
)

rds <- file.path(cache_dir, paste(sub_name, names(timepoints)[2], "markers.rds", sep = "_"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(timepoints[[2]], only.pos = TRUE, verbose = FALSE)
  rownames(markers) <- NULL
  saveRDS(markers, rds)
}

datatable_download(markers)
```

# IPC

```{r}
sub_name <- "IPC"
```

## D72

```{r}
timepoints[[1]] <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, names(timepoints)[1], "seurat.rds", sep = "_"))
)

rds <- file.path(cache_dir, paste(sub_name, names(timepoints)[1], "markers.rds", sep = "_"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(timepoints[[1]], only.pos = TRUE, verbose = FALSE)
  rownames(markers) <- NULL
  saveRDS(markers, rds)
}

datatable_download(markers)
```

## D120

```{r}
timepoints[[2]] <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, names(timepoints)[2], "seurat.rds", sep = "_"))
)

rds <- file.path(cache_dir, paste(sub_name, names(timepoints)[2], "markers.rds", sep = "_"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(timepoints[[2]], only.pos = TRUE, verbose = FALSE)
  rownames(markers) <- NULL
  saveRDS(markers, rds)
}

datatable_download(markers)
```

# Mural

```{r}
sub_name <- "Mu"
```

## D72

```{r}
timepoints[[1]] <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, names(timepoints)[1], "seurat.rds", sep = "_"))
)

rds <- file.path(cache_dir, paste(sub_name, names(timepoints)[1], "markers.rds", sep = "_"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(timepoints[[1]], only.pos = TRUE, verbose = FALSE)
  rownames(markers) <- NULL
  saveRDS(markers, rds)
}

datatable_download(markers)
```

## D120

```{r}
timepoints[[2]] <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, names(timepoints)[2], "seurat.rds", sep = "_"))
)

rds <- file.path(cache_dir, paste(sub_name, names(timepoints)[2], "markers.rds", sep = "_"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(timepoints[[2]], only.pos = TRUE, verbose = FALSE)
  rownames(markers) <- NULL
  saveRDS(markers, rds)
}

datatable_download(markers)
```

# Radial Glia

```{r}
sub_name <- "RG"
```

## D72

```{r}
timepoints[[1]] <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, names(timepoints)[1], "seurat.rds", sep = "_"))
)

rds <- file.path(cache_dir, paste(sub_name, names(timepoints)[1], "markers.rds", sep = "_"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(timepoints[[1]], only.pos = TRUE, verbose = FALSE)
  rownames(markers) <- NULL
  saveRDS(markers, rds)
}

datatable_download(markers)
```

## D120

```{r}
timepoints[[2]] <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, names(timepoints)[2], "seurat.rds", sep = "_"))
)

rds <- file.path(cache_dir, paste(sub_name, names(timepoints)[2], "markers.rds", sep = "_"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(timepoints[[2]], only.pos = TRUE, verbose = FALSE)
  rownames(markers) <- NULL
  saveRDS(markers, rds)
}

datatable_download(markers)
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
