---
title: "06 WGCNA - Combined Organelles"
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "Diana Acosta [Diana.Acosta@osumc.edu]"
  - name: "Hongjun Fu [Hongjun.Fu@osumc.edu]"
  - name: "Ohio State University Wexner Medical Center"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[8], "06_WGCNA.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Acosta Organoid scRNAseq Project](https://github.com/eturkes/acosta-organoid-scRNAseq).*

In this document we perform WGCNA analysis.
The data here will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[8]``.

```{r}
#    This file is part of acosta-organoid-scRNAseq.
#    Copyright (C) 2022-2025  Emir Turkes, Diana Acosta, Gi Guo, Yuzhou Chang,
#    Hongjun Fu, Ohio State University Wexner Medical Center
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

packages <- c("conflicted", "Seurat", "WGCNA", "flashClust", "GSVA", "ggplot2")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path("..", "utils.R"))

vars_to_regress <- NULL # See `cluster_pipeline` in `utils.R`.
cex1 <- 0.9

analysis_no <- 6
data_name <- unlist(strsplit(getwd(), "/"))[8] # Name of dataset.
data_dir <- file.path("..", "..", "data") # Backed up data.
results_dir <- file.path("..", "..", "results")

# Unique cache directory for each analysis number.
# ------------------------------------------------
cache_dir <- file.path("..", "..", "cache", data_name, paste0("0", analysis_no))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# ------------------------------------------------

knitr::opts_chunk$set(fig.width = 12, fig.height = 7, dpi = 300)
```

# Prep

```{r}
seurat <- readRDS(file.path(cache_dir, "..", "02", "rca_seurat.rds"))

# Adapted from code written by Gi Guo and Yuzhou Chang.
# -----------------------------------------------------
rds <- file.path(cache_dir, "All_HVG.rds")
if (file.exists(rds)) {
  All_HVG <- readRDS(rds)
} else {

  All_HVG <- vector("list", length(unique(seurat$sample)) + 1)
  for (i in seq_along(All_HVG)) {

    if (i == length(unique(seurat$sample)) + 1) {
      sub_seurat <- seurat
    } else {
      sub_seurat <- seurat[ , which(seurat$sample == unique(seurat$sample)[i])]
    }

    sub_seurat <- suppressWarnings(
      SCTransform(
        sub_seurat, vars.to.regress = vars_to_regress, vst.flavor = "v2",
        return.only.var.genes = FALSE, verbose = FALSE
      )
    )
    HVG <- FindVariableFeatures(
      sub_seurat,
      selection.method = "vst",
      nfeatures = 10000,
      assay = "SCT"
    )
    All_HVG[[i]] <- HVG@assays$SCT@var.features
  }

  rm(HVG, sub_seurat)
  All_HVG <- unique(unlist(All_HVG))
  saveRDS(All_HVG, rds, compress = FALSE)
}

all_matrix_hvg <- seurat@assays$SCT@data
all_matrix_hvg <- seurat@assays$SCT@data[match(All_HVG, rownames(all_matrix_hvg)), ]
all_matrix_hvg <- t(as.matrix(all_matrix_hvg))
gene.names <- colnames(all_matrix_hvg)
# -----------------------------------------------------
```

# WGCNA Network Creation

```{r}
# Adapted from code written by Gi Guo and Yuzhou Chang.
# -----------------------------------------------------
# Calculate soft threshold-β
# --------------------------
powers <- c(c(1:10), seq(from = 12, to = 20, by = 2)) # Better to include powers up to 20.

rds <- file.path(cache_dir, "sft.rds")
if (file.exists(rds)) {
  sft <- readRDS(rds)
} else {
  sft <- pickSoftThreshold(
    all_matrix_hvg,
    dataIsExpr = TRUE,
    powerVector = powers,
    corFnc = cor,
    corOptions = list(use = 'p'),
    networkType = "unsigned"
  )
  saveRDS(sft, rds, compress = FALSE)
}
# --------------------------

# SFT index as a function of different powers.
# --------------------------------------------
plot(
  sft$fitIndices[, 1],
  -sign(sft$fitIndices[, 3]) * sft$fitIndices[, 2],
  xlab = "Soft Threshold (power)",
  ylab = "Scale Free Topology Model Fit, signed R^2",
  type = "n",
  main = paste("Scale independence")
)
text(
  sft$fitIndices[, 1],
  -sign(sft$fitIndices[, 3]) * sft$fitIndices[, 2],
  labels = powers,
  cex = cex1,
  col = "red"
)
abline(h = 0.9, col = "red") # This line corresponds to using an R^2 cut-off of h.
# --------------------------------------------

# Mean connectivity as a function of different powers.
# ----------------------------------------------------
plot(
  sft$fitIndices[, 1],
  sft$fitIndices[, 5],
  xlab = "Soft Threshold (power)",
  ylab = "Mean Connectivity",
  type = "n",
  main = paste("Mean connectivity")
)
text(
  sft$fitIndices[, 1],
  sft$fitIndices[, 5],
  labels = powers,
  cex = cex1,
  col = "red"
)
# ----------------------------------------------------

# Automatic module detection.
# ---------------------------
softPower <- 2

rds <- file.path(cache_dir, "bwnet.rds")
if (file.exists(rds)) {
  bwnet <- readRDS(rds)
} else {
  bwnet <- blockwiseModules(
    all_matrix_hvg,
    maxBlockSize = 5000,
    mergeCutHeight = 0.25,
    networkType = "unsigned",
    power = softPower,
    minModuleSize = 30,
    numericLabels = TRUE,
    saveTOMs = FALSE,
    pamRespectsDendro = FALSE,
    verbose = 3
  )
  saveRDS(bwnet, rds, compress = FALSE)
}
# ---------------------------

rds <- file.path(cache_dir, "TOM.rds")
if (file.exists(rds)) {
  TOM <- readRDS(rds)
} else {
  TOM <- TOMsimilarityFromExpr(
    all_matrix_hvg,
    networkType = "unsigned",
    TOMType = "unsigned",
    power = softPower
  )
  colnames(TOM) <- rownames(TOM) <- gene.names
  saveRDS(TOM, rds, compress = FALSE)
}

rds <- file.path(cache_dir, "geneTree.rds")
if (file.exists(rds)) {
  geneTree <- readRDS(rds)
} else {
  dissTOM <- 1 - TOM
  geneTree <- flashClust(as.dist(dissTOM), method = "average")
  rm(dissTOM)
  saveRDS(geneTree, rds, compress = FALSE)
}

plot(geneTree, xlab = "", sub = "", cex = 0.3)

# Module identification using dynamic tree cut.
# ---------------------------------------------
dynamicMods <- cutreeDynamic(
  dendro = geneTree,
  method = "tree",
  minClusterSize = 20,
  deepSplit = 3
)
table(dynamicMods)
# ---------------------------------------------

# Plot the module assignment under the dendrogram.
# Note: The grey color is reserved for unassigned genes.
# ------------------------------------------------------
dynamicColors <- labels2colors(dynamicMods)
table(dynamicColors)

plotDendroAndColors(
  geneTree,
  dynamicColors,
  "Dynamic Tree Cut",
  dendroLabels = FALSE,
  hang = 0.03,
  addGuide = TRUE,
  guideHang = 0.05,
  main = "Gene dendrogram and module colors"
)
# ------------------------------------------------------

# Discard the unassigned genes, and focus on the rest.
# ----------------------------------------------------
restGenes <- dynamicColors != "grey"

rds <- file.path(cache_dir, "diss1.rds")
if (file.exists(rds)) {
  diss1 <- readRDS(rds)
} else {
  diss1 <- 1 - TOMsimilarityFromExpr(all_matrix_hvg[ , restGenes], power = softPower)
  colnames(diss1) <- rownames(diss1) <- gene.names[restGenes]
  saveRDS(diss1, rds, compress = FALSE)
}
# ----------------------------------------------------

hier1 <- flashClust(as.dist(diss1), method = "average")
plotDendroAndColors(
  hier1,
  dynamicColors[restGenes],
  "Dynamic Tree Cut",
  dendroLabels = FALSE,
  hang = 0.03,
  addGuide = TRUE,
  guideHang = 0.05,
  main = "Gene dendrogram and module colors"
)

diag(diss1) <- NA # Set the diagonal of the dissimilarity to NA.

# Visualize the Tom plot.
# Raise the dissimilarity matrix to the power of 4 to bring out the module structure.
# -----------------------------------------------------------------------------------
TOMplot(diss1, hier1, as.character(dynamicColors[restGenes]))
# -----------------------------------------------------------------------------------

# Extract module.
# ---------------
module_colors <- setdiff(unique(dynamicColors), "grey")
module_list <- vector("list", length(module_colors))
names(module_list) <- module_colors
for (i in seq_along(module_list)) {
  module_list[[i]] <- gene.names[which(dynamicColors == module_colors[i])]
}

data <- data.frame(1:8)
for (i in 1:length(module_colors)) {
  data[i, 1] <- length(gene.names[which(dynamicColors == module_colors[i])])
}
rownames(data) <- module_colors
colnames(data) <- c("gene_num")
# ---------------
# -----------------------------------------------------
```

# Analysis

```{r}
rds <- file.path(cache_dir, "gsva.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(
    GetAssayData(seurat), module_list, method = "ssgsea", ssgsea.norm = FALSE, verbose = FALSE
  )
  saveRDS(gsva, rds, compress = FALSE)
}
gsva <- (2 * (gsva - min(gsva)) / (max(gsva) - min(gsva))) - 1

seurat[["GSVA"]] <- CreateAssayObject(data = as.matrix(gsva))
DefaultAssay(seurat) <- "GSVA"

for (i in seq(nrow(seurat))) {
  print(
    FeaturePlot(seurat, rownames(seurat)[i], order = TRUE, reduction = "rca_umap") +
      scale_color_gradientn(colors = c("blue", "#F8F8F8", "red"),  limits = c(-1, 1))
  )
  print(
    RidgePlot(seurat, rownames(seurat)[i])
  )
  print(
    RidgePlot(seurat, rownames(seurat)[i], group.by = "genotype")
  )
}

DotPlot(seurat, features = rownames(seurat), cols = c("blue", "red")) + RotatedAxis()
DotPlot(seurat, features = rownames(seurat), cols = c("blue", "red"), group.by = "genotype") +
  RotatedAxis()
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
