---
title: "10 Differential Abundance - Combined Organelles"
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "Diana Acosta [Diana.Acosta@osumc.edu]"
  - name: "Hongjun Fu [Hongjun.Fu@osumc.edu]"
  - name: "Ohio State University Wexner Medical Center"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[8], "10_abundance.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
.tocify-subheader .tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 45px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Acosta Organoid scRNAseq Project](https://github.com/eturkes/acosta-organoid-scRNAseq).*

In this document we perform differential gene expression analysis.
The data here will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[8]``.

```{r}
#    This file is part of acosta-organoid-scRNAseq.
#    Copyright (C) 2022  Emir Turkes, Diana Acosta, Gi Guo, Yuzhou Chang,
#    Hongjun Fu, Ohio State University Wexner Medical Center
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

packages <- c("conflicted", "Seurat", "ggplot2")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path("..", "utils.R"))

analysis_no <- 10
data_name <- unlist(strsplit(getwd(), "/"))[8] # Name of dataset.
data_dir <- file.path("..", "..", "data") # Backed up data.
results_dir <- file.path("..", "..", "results")

# Unique cache directory for each analysis number.
# ------------------------------------------------
cache_dir <- file.path("..", "..", "cache", data_name, paste0("0", analysis_no))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# ------------------------------------------------

knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
```

# All

```{r}
seurat_list <- readRDS(file.path(cache_dir, "..", "03", "seurat.rds"))
timepoint_seurat <- SplitObject(seurat_list, "timepoint")
seurat_list <- c(D120 = timepoint_seurat$D120, D72 = timepoint_seurat$D72, comb = seurat_list)
rm(timepoint_seurat)

for (i in seq_along(seurat_list)) {
  seurat_list[[i]]$genotype <- factor(seurat_list[[i]]$genotype, c("WT", "HET", "HOM"))
  seurat_list[[i]]$timepoint <- factor(seurat_list[[i]]$timepoint, c("D72", "D120"))
}

df <- as.data.frame(table(seurat_list$comb$subclusters, seurat_list$comb$timepoint))
colnames(df) <- c("Celltype", "Timepoint", "Proportion")
ggplot(df, aes(Timepoint, Proportion, fill = Celltype)) +
  geom_bar(position = "fill", stat = "identity")

df <- as.data.frame(table(seurat_list$comb$subclusters, seurat_list$comb$timepoint))
colnames(df) <- c("Celltype", "Timepoint", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Timepoint)) +
  geom_bar(position = "fill", stat = "identity") +
  theme(axis.text.x = element_text(angle = 90))

df <- as.data.frame(table(seurat_list$D120$subclusters, seurat_list$D120$genotype))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Genotype)) +
  geom_bar(position = "fill", stat = "identity") +
  theme(axis.text.x = element_text(angle = 90))

df <- as.data.frame(table(seurat_list$D72$subclusters, seurat_list$D72$genotype))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Genotype)) +
  geom_bar(position = "fill", stat = "identity") +
  theme(axis.text.x = element_text(angle = 90))

df <- as.data.frame(table(seurat_list$comb$cluster_names, seurat_list$comb$timepoint))
colnames(df) <- c("Celltype", "Timepoint", "Proportion")
ggplot(df, aes(Timepoint, Proportion, fill = Celltype)) +
  geom_bar(position = "fill", stat = "identity")

df <- as.data.frame(table(seurat_list$comb$cluster_names, seurat_list$comb$timepoint))
colnames(df) <- c("Celltype", "Timepoint", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Timepoint)) +
  geom_bar(position = "fill", stat = "identity")

df <- as.data.frame(table(seurat_list$D120$cluster_names, seurat_list$D120$genotype))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Genotype)) +
  geom_bar(position = "fill", stat = "identity")

df <- as.data.frame(table(seurat_list$D72$cluster_names, seurat_list$D72$genotype))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Genotype)) +
  geom_bar(position = "fill", stat = "identity")
```

# Excitatory Neuron

```{r}
sub_name <- "Ex"
seurat_list <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, "seurat_list.rds", sep = "_"))
)

for (i in seq_along(seurat_list)) {
  seurat_list[[i]]$genotype <- factor(seurat_list[[i]]$genotype, c("WT", "HET", "HOM"))
  seurat_list[[i]]$timepoint <- factor(seurat_list[[i]]$timepoint, c("D72", "D120"))
}

df <- as.data.frame(table(seurat_list$comb$clustermap_names_short, seurat_list$comb$timepoint))
colnames(df) <- c("Celltype", "Timepoint", "Proportion")
ggplot(df, aes(Timepoint, Proportion, fill = Celltype)) +
  geom_bar(position = "fill", stat = "identity")

df <- as.data.frame(table(seurat_list$comb$clustermap_names_short, seurat_list$comb$timepoint))
colnames(df) <- c("Celltype", "Timepoint", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Timepoint)) +
  geom_bar(position = "fill", stat = "identity")

df <- as.data.frame(table(seurat_list$D120$clustermap_names_short, seurat_list$D120$genotype))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Genotype)) +
  geom_bar(position = "fill", stat = "identity")

df <- as.data.frame(table(seurat_list$D72$clustermap_names_short, seurat_list$D72$genotype))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Genotype)) +
  geom_bar(position = "fill", stat = "identity")
```

# Radial Glia

```{r}
sub_name <- "RG"
seurat_list <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, "seurat_list.rds", sep = "_"))
)

for (i in seq_along(seurat_list)) {
  seurat_list[[i]]$genotype <- factor(seurat_list[[i]]$genotype, c("WT", "HET", "HOM"))
  seurat_list[[i]]$timepoint <- factor(seurat_list[[i]]$timepoint, c("D72", "D120"))
}

df <- as.data.frame(table(seurat_list$comb$clustermap_names_short, seurat_list$comb$timepoint))
colnames(df) <- c("Celltype", "Timepoint", "Proportion")
ggplot(df, aes(Timepoint, Proportion, fill = Celltype)) +
  geom_bar(position = "fill", stat = "identity")

df <- as.data.frame(table(seurat_list$comb$clustermap_names_short, seurat_list$comb$timepoint))
colnames(df) <- c("Celltype", "Timepoint", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Timepoint)) +
  geom_bar(position = "fill", stat = "identity")

df <- as.data.frame(table(seurat_list$D120$clustermap_names_short, seurat_list$D120$genotype))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Genotype)) +
  geom_bar(position = "fill", stat = "identity")

df <- as.data.frame(table(seurat_list$D72$clustermap_names_short, seurat_list$D72$genotype))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Genotype)) +
  geom_bar(position = "fill", stat = "identity")
```

# IPC

```{r}
sub_name <- "IPC"
seurat_list <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, "seurat_list.rds", sep = "_"))
)

for (i in seq_along(seurat_list)) {
  seurat_list[[i]]$genotype <- factor(seurat_list[[i]]$genotype, c("WT", "HET", "HOM"))
  seurat_list[[i]]$timepoint <- factor(seurat_list[[i]]$timepoint, c("D72", "D120"))
}

df <- as.data.frame(table(seurat_list$comb$clustermap_names_short, seurat_list$comb$timepoint))
colnames(df) <- c("Celltype", "Timepoint", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Timepoint)) +
  geom_bar(position = "fill", stat = "identity")

df <- as.data.frame(table(seurat_list$D120$clustermap_names_short, seurat_list$D120$genotype))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Genotype)) +
  geom_bar(position = "fill", stat = "identity")

df <- as.data.frame(table(seurat_list$D72$clustermap_names_short, seurat_list$D72$genotype))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Genotype)) +
  geom_bar(position = "fill", stat = "identity")
```

# Inhibitory Neuron

```{r}
sub_name <- "In"
seurat_list <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, "seurat_list.rds", sep = "_"))
)

for (i in seq_along(seurat_list)) {
  seurat_list[[i]]$genotype <- factor(seurat_list[[i]]$genotype, c("WT", "HET", "HOM"))
  seurat_list[[i]]$timepoint <- factor(seurat_list[[i]]$timepoint, c("D72", "D120"))
}

df <- as.data.frame(table(seurat_list$comb$clustermap_names_short, seurat_list$comb$timepoint))
colnames(df) <- c("Celltype", "Timepoint", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Timepoint)) +
  geom_bar(position = "fill", stat = "identity")

df <- as.data.frame(table(seurat_list$D120$clustermap_names_short, seurat_list$D120$genotype))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Genotype)) +
  geom_bar(position = "fill", stat = "identity")

df <- as.data.frame(table(seurat_list$D72$clustermap_names_short, seurat_list$D72$genotype))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Genotype)) +
  geom_bar(position = "fill", stat = "identity")
```

# Mural

```{r}
sub_name <- "Mu"
seurat_list <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, "seurat_list.rds", sep = "_"))
)

for (i in seq_along(seurat_list)) {
  seurat_list[[i]]$genotype <- factor(seurat_list[[i]]$genotype, c("WT", "HET", "HOM"))
  seurat_list[[i]]$timepoint <- factor(seurat_list[[i]]$timepoint, c("D72", "D120"))
}

df <- as.data.frame(table(seurat_list$comb$clustermap_names_short, seurat_list$comb$timepoint))
colnames(df) <- c("Celltype", "Timepoint", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Timepoint)) +
  geom_bar(position = "fill", stat = "identity")

df <- as.data.frame(table(seurat_list$D120$clustermap_names_short, seurat_list$D120$genotype))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Genotype)) +
  geom_bar(position = "fill", stat = "identity")

df <- as.data.frame(table(seurat_list$D72$clustermap_names_short, seurat_list$D72$genotype))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Genotype)) +
  geom_bar(position = "fill", stat = "identity")
```

# Unclassified

```{r}
sub_name <- "Unc"
seurat_list <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, "seurat_list.rds", sep = "_"))
)

for (i in seq_along(seurat_list)) {
  seurat_list[[i]]$genotype <- factor(seurat_list[[i]]$genotype, c("WT", "HET", "HOM"))
  seurat_list[[i]]$timepoint <- factor(seurat_list[[i]]$timepoint, c("D72", "D120"))
}

df <- as.data.frame(table(seurat_list$comb$clustermap_names_short, seurat_list$comb$timepoint))
colnames(df) <- c("Celltype", "Timepoint", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Timepoint)) +
  geom_bar(position = "fill", stat = "identity")

df <- as.data.frame(table(seurat_list$D120$clustermap_names_short, seurat_list$D120$genotype))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Genotype)) +
  geom_bar(position = "fill", stat = "identity")

df <- as.data.frame(table(seurat_list$D72$clustermap_names_short, seurat_list$D72$genotype))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Celltype, Proportion, fill = Genotype)) +
  geom_bar(position = "fill", stat = "identity")
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
