---
title: "12 Markers - Combined Organelles"
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "Diana Acosta [Diana.Acosta@osumc.edu]"
  - name: "Hongjun Fu [Hongjun.Fu@osumc.edu]"
  - name: "Ohio State University Wexner Medical Center"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[8], "12_markers.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Acosta Organoid scRNAseq Project](https://github.com/eturkes/acosta-organoid-scRNAseq).*

In this document we annotate cell-types based on reference datasets and marker genes.
The data here will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[8]``.

```{r}
#    This file is part of acosta-organoid-scRNAseq.
#    Copyright (C) 2022-2024  Emir Turkes, Diana Acosta, Gi Guo, Yuzhou Chang,
#    Hongjun Fu, Ohio State University Wexner Medical Center
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

packages <- c(
  "conflicted", "Seurat", "magrittr", "dplyr", "ComplexHeatmap", "SingleCellExperiment", "scuttle",
  "edgeR", "DT"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path("..", "utils.R"))
conflict_prefer("cpm", "edgeR", quiet = TRUE)
conflict_prefer("slice", "dplyr")

analysis_no <- 12
data_name <- unlist(strsplit(getwd(), "/"))[8] # Name of dataset.
data_dir <- file.path("..", "..", "data") # Backed up data.
results_dir <- file.path("..", "..", "results")

# Unique cache directory for each analysis number.
# ------------------------------------------------
cache_dir <- file.path("..", "..", "cache", data_name, analysis_no)
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# ------------------------------------------------

knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
```

# All

```{r}
seurat <- readRDS(file.path(cache_dir, "..", "03", "seurat.rds"))

rds <- file.path(cache_dir, "broad_markers.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds, compress = FALSE)
}
markers$cluster <- factor(markers$cluster, sort(unique(as.character(markers$cluster))))
markers <- markers[markers$p_val_adj < 0.05, ]
datatable_download(markers)
markers <- markers %>% group_by(cluster) %>% slice_head(n = 5)

sce <- SingleCellExperiment(list(counts = GetAssayData(seurat, slot = "counts")))
sce$cluster_names <- seurat$cluster_names
sce <- suppressWarnings(aggregateAcrossCells(sce, sce$cluster_names))
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

sce_sub <- sce[rownames(sce) %in% markers$gene, ]
sce_sub <- sce_sub[match(markers$gene, rownames(sce_sub)), ]

mat <- t(apply(logcounts(sce_sub), 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    column_order = c(2, 5, 1, 4, 6, 3),
    heatmap_legend_param = list(
      title = "Per-row Scaled Aggregated Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)
```

# Excitatory Neuron

```{r}
sub_name <- "Ex"
seurat <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, "seurat_list.rds", sep = "_"))
)
seurat <- seurat$comb

rds <- file.path(cache_dir, paste(sub_name, "subcluster", "markers.rds", sep = "_"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds, compress = FALSE)
}
markers$cluster <- factor(markers$cluster, sort(unique(as.character(markers$cluster))))
markers <- markers[markers$p_val_adj < 0.05, ]
datatable_download(markers)
markers <- markers %>% group_by(cluster) %>% slice_head(n = 1)

sce <- SingleCellExperiment(list(counts = GetAssayData(seurat, slot = "counts")))
sce$clustermap_names <- seurat$clustermap_names
sce <- suppressWarnings(aggregateAcrossCells(sce, sce$clustermap_names))
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)
sce <- sce[rownames(sce) %in% markers$gene, ]
sce <- sce[match(markers$gene, rownames(sce)), ]

mat <- t(apply(logcounts(sce), 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    heatmap_legend_param = list(
      title = "Per-row Scaled Aggregated Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)

seurat@active.ident <- seurat$clustermap_names_short

rds <- file.path(cache_dir, paste(sub_name, "subcluster_short", "markers.rds", sep = "_"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds, compress = FALSE)
}
markers$cluster <- factor(markers$cluster, sort(unique(as.character(markers$cluster))))
markers <- markers[markers$p_val_adj < 0.05, ]
datatable_download(markers)
markers <- markers %>% group_by(cluster) %>% slice_head(n = 5)

sce <- SingleCellExperiment(list(counts = GetAssayData(seurat, slot = "counts")))
sce$clustermap_names_short <- seurat$clustermap_names_short
sce <- suppressWarnings(aggregateAcrossCells(sce, sce$clustermap_names_short))
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)
sce <- sce[rownames(sce) %in% markers$gene, ]
sce <- sce[match(markers$gene, rownames(sce)), ]

mat <- t(apply(logcounts(sce), 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    column_names_rot = 45,
    heatmap_legend_param = list(
      title = "Per-row Scaled Aggregated Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)

seurat <- seurat[ , seurat$clustermap_names_short == "Ex Upper Layer"]
seurat@active.ident <- seurat$clustermap_names

rds <- file.path(cache_dir, paste(sub_name, "upper_layer", "markers.rds", sep = "_"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds, compress = FALSE)
}
markers$cluster <- factor(markers$cluster, sort(unique(as.character(markers$cluster))))
markers <- markers[markers$p_val_adj < 0.05, ]
datatable_download(markers)
markers <- markers %>% group_by(cluster) %>% slice_head(n = 5)

sce <- SingleCellExperiment(list(counts = GetAssayData(seurat, slot = "counts")))
sce$clustermap_names <- seurat$clustermap_names
sce <- suppressWarnings(aggregateAcrossCells(sce, sce$clustermap_names))
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)
sce <- sce[rownames(sce) %in% markers$gene, ]
sce <- sce[match(markers$gene, rownames(sce)), ]

mat <- t(apply(logcounts(sce), 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    heatmap_legend_param = list(
      title = "Per-row Scaled Aggregated Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)
```

# Radial Glia

```{r, fig.width = 12, fig.height = 10}
sub_name <- "RG"
seurat <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, "seurat_list.rds", sep = "_"))
)
seurat <- seurat$comb

rds <- file.path(cache_dir, paste(sub_name, "subcluster", "markers.rds", sep = "_"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds, compress = FALSE)
}
markers$cluster <- factor(markers$cluster, sort(unique(as.character(markers$cluster))))
markers <- markers[markers$p_val_adj < 0.05, ]
datatable_download(markers)
markers <- markers %>% group_by(cluster) %>% slice_head(n = 1)

sce <- SingleCellExperiment(list(counts = GetAssayData(seurat, slot = "counts")))
sce$clustermap_names <- seurat$clustermap_names
sce <- suppressWarnings(aggregateAcrossCells(sce, sce$clustermap_names))
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)
sce <- sce[rownames(sce) %in% markers$gene, ]
sce <- sce[match(markers$gene, rownames(sce)), ]

mat <- t(apply(logcounts(sce), 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    heatmap_legend_param = list(
      title = "Per-row Scaled Aggregated Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)

seurat@active.ident <- seurat$clustermap_names_short

rds <- file.path(cache_dir, paste(sub_name, "subcluster_short", "markers.rds", sep = "_"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds, compress = FALSE)
}
markers$cluster <- factor(markers$cluster, sort(unique(as.character(markers$cluster))))
markers <- markers[markers$p_val_adj < 0.05, ]
datatable_download(markers)
markers <- markers %>% group_by(cluster) %>% slice_head(n = 5)

sce <- SingleCellExperiment(list(counts = GetAssayData(seurat, slot = "counts")))
sce$clustermap_names_short <- seurat$clustermap_names_short
sce <- suppressWarnings(aggregateAcrossCells(sce, sce$clustermap_names_short))
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)
sce <- sce[rownames(sce) %in% markers$gene, ]
sce <- sce[match(markers$gene, rownames(sce)), ]

mat <- t(apply(logcounts(sce), 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    heatmap_legend_param = list(
      title = "Per-row Scaled Aggregated Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)
```

# IPC

```{r}
sub_name <- "IPC"
seurat <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, "seurat_list.rds", sep = "_"))
)
seurat <- seurat$comb

rds <- file.path(cache_dir, paste(sub_name, "subcluster", "markers.rds", sep = "_"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds, compress = FALSE)
}
markers$cluster <- factor(markers$cluster, sort(unique(as.character(markers$cluster))))
markers <- markers[markers$p_val_adj < 0.05, ]
datatable_download(markers)
markers <- markers %>% group_by(cluster) %>% slice_head(n = 1)

sce <- SingleCellExperiment(list(counts = GetAssayData(seurat, slot = "counts")))
sce$clustermap_names <- seurat$clustermap_names
sce <- suppressWarnings(aggregateAcrossCells(sce, sce$clustermap_names))
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)
sce <- sce[rownames(sce) %in% markers$gene, ]
sce <- sce[match(markers$gene, rownames(sce)), ]

mat <- t(apply(logcounts(sce), 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    heatmap_legend_param = list(
      title = "Per-row Scaled Aggregated Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)
```

# Inhibitory Neuron

```{r}
sub_name <- "In"
seurat <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, "seurat_list.rds", sep = "_"))
)
seurat <- seurat$comb

rds <- file.path(cache_dir, paste(sub_name, "subcluster", "markers.rds", sep = "_"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds, compress = FALSE)
}
markers$cluster <- factor(markers$cluster, sort(unique(as.character(markers$cluster))))
markers <- markers[markers$p_val_adj < 0.05, ]
datatable_download(markers)
markers <- markers %>% group_by(cluster) %>% slice_head(n = 1)

sce <- SingleCellExperiment(list(counts = GetAssayData(seurat, slot = "counts")))
sce$clustermap_names <- seurat$clustermap_names
sce <- suppressWarnings(aggregateAcrossCells(sce, sce$clustermap_names))
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)
sce <- sce[rownames(sce) %in% markers$gene, ]
sce <- sce[match(markers$gene, rownames(sce)), ]

mat <- t(apply(logcounts(sce), 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    heatmap_legend_param = list(
      title = "Per-row Scaled Aggregated Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)

seurat@active.ident <- seurat$clustermap_names_short

rds <- file.path(cache_dir, paste(sub_name, "subcluster_short", "markers.rds", sep = "_"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds, compress = FALSE)
}
markers$cluster <- factor(markers$cluster, sort(unique(as.character(markers$cluster))))
markers <- markers[markers$p_val_adj < 0.05, ]
datatable_download(markers)
markers <- markers %>% group_by(cluster) %>% slice_head(n = 5)

sce <- SingleCellExperiment(list(counts = GetAssayData(seurat, slot = "counts")))
sce$clustermap_names_short <- seurat$clustermap_names_short
sce <- suppressWarnings(aggregateAcrossCells(sce, sce$clustermap_names_short))
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)
sce <- sce[rownames(sce) %in% markers$gene, ]
sce <- sce[match(markers$gene, rownames(sce)), ]

mat <- t(apply(logcounts(sce), 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    heatmap_legend_param = list(
      title = "Per-row Scaled Aggregated Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)
```

# Mural

```{r}
sub_name <- "Mu"
seurat <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, "seurat_list.rds", sep = "_"))
)
seurat <- seurat$comb

rds <- file.path(cache_dir, paste(sub_name, "subcluster", "markers.rds", sep = "_"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds, compress = FALSE)
}
markers$cluster <- factor(markers$cluster, sort(unique(as.character(markers$cluster))))
markers <- markers[markers$p_val_adj < 0.05, ]
datatable_download(markers)
markers <- markers %>% group_by(cluster) %>% slice_head(n = 1)

sce <- SingleCellExperiment(list(counts = GetAssayData(seurat, slot = "counts")))
sce$clustermap_names <- seurat$clustermap_names
sce <- suppressWarnings(aggregateAcrossCells(sce, sce$clustermap_names))
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)
sce <- sce[rownames(sce) %in% markers$gene, ]
sce <- sce[match(markers$gene, rownames(sce)), ]

mat <- t(apply(logcounts(sce), 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    heatmap_legend_param = list(
      title = "Per-row Scaled Aggregated Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)
```

# Unclassified

```{r}
sub_name <- "Unc"
seurat <- readRDS(
  file.path(cache_dir, "..", "03", paste(sub_name, "seurat_list.rds", sep = "_"))
)
seurat <- seurat$comb

rds <- file.path(cache_dir, paste(sub_name, "subcluster", "markers.rds", sep = "_"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds, compress = FALSE)
}
markers$cluster <- factor(markers$cluster, sort(unique(as.character(markers$cluster))))
markers <- markers[markers$p_val_adj < 0.05, ]
datatable_download(markers)
markers <- markers %>% group_by(cluster) %>% slice_head(n = 1)

sce <- SingleCellExperiment(list(counts = GetAssayData(seurat, slot = "counts")))
sce$clustermap_names <- seurat$clustermap_names
sce <- suppressWarnings(aggregateAcrossCells(sce, sce$clustermap_names))
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)
sce <- sce[rownames(sce) %in% markers$gene, ]
sce <- sce[match(markers$gene, rownames(sce)), ]

mat <- t(apply(logcounts(sce), 1, function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)))
set.seed(1)
draw(
  Heatmap(
    mat,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    heatmap_legend_param = list(
      title = "Per-row Scaled Aggregated Expression", direction = "horizontal",
      legend_width = unit(7.5, "cm")
    )
  ),
  heatmap_legend_side = "top"
)
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
