---
title: "03 Subclusters - Glasauer et al. 2022"
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "Diana Acosta [Diana.Acosta@osumc.edu]"
  - name: "Hongjun Fu [Hongjun.Fu@osumc.edu]"
  - name: "Ohio State University Wexner Medical Center"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[8], "03_subclusters.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Acosta Organoid scRNAseq Project](https://github.com/eturkes/acosta-organoid-scRNAseq).*

In this document we subcluster annotated cell-types.
The data here will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[8]``.

```{r}
#    This file is part of acosta-organoid-scRNAseq.
#    Copyright (C) 2022  Emir Turkes, Diana Acosta, Gi Guo, Yuzhou Chang,
#    Hongjun Fu, Ohio State University Wexner Medical Center
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

packages <- c(
  "conflicted", "Seurat", "ggplot2", "dplyr", "ggrepel", "SingleCellExperiment", "scuttle", "RCAv2",
  "data.table", "GSVA", "edgeR", "ComplexHeatmap", "ClusterMap", "filesstrings", "patchwork", "DT",
  "viridis"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
options(future.globals.maxSize = 10485760000)
source(file.path("..", "utils.R"))

protocol <- c("human", "droplet", "single-cell", "umis") # See `cluster_pipeline` in `utils.R`.
vars_to_regress <- NULL # See `cluster_pipeline`.
parallel_override <- 0 # See `parallel_plan` in `utils.R`.

# Metadata to plot after dimensionality reduction and clustering.
# Values in list can include "no_legend and/or "no_label" to exclude those.
# -------------------------------------------------------------------------
metadata_to_plot <- vector("list", 4)
names(metadata_to_plot) <- c("seurat_clusters", "sample", "timepoint", "genotype")
# -------------------------------------------------------------------------

analysis_no <- 3
data_name <- unlist(strsplit(getwd(), "/"))[8] # Name of dataset.
data_dir <- file.path("..", "..", "data") # Backed up data.
results_dir <- file.path("..", "..", "results")

# Unique cache directory for each analysis number.
# ------------------------------------------------
cache_dir <- file.path("..", "..", "cache", data_name, paste0("0", analysis_no))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# ------------------------------------------------

knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 150)
```

# Prep

```{r}
seurat <- readRDS(file.path(cache_dir, "..", "02", "rca_seurat.rds"))
seurat

red_dim_plot(seurat, "rca_umap1", "rca_umap2", "cluster_names", "cat")

markers <- read.delim(
  file.path(data_dir, "gene-lists", "Human cortical development-gene-list summary - page3.tsv"),
  check.names = FALSE
)

markers <- markers[ , 1:which(colnames(markers) == "Polioudakis Neuron 2019")]
markers <- markers[ , -(which(colnames(markers) == "oRG vRG shared genes"))]
markers <- Filter(function(x) !all(is.na(x)), markers)
markers <- as.list(markers)
markers <- lapply(markers, function(x) x[!(x == "")])

remove <- c(
  "SORCS1", "ADRA2A", "SLA", "STMN2", "SORBS2", "ELAVL2", "DCX", "ZFHX4", "GLI3", "CD99", "IGFBP7",
  "AURKC", "HMGB1", "HMGB3", "CTNNB1", "HES1", "VIM", "PTN", "CLU", "IL33", "NPR3", "CCNB2",
  "KIF15", "ARHGAP11B", "CENPK", "NEUROD2", "NEUROD6", "DLX2", "ERBB4", "SOX5", "BCL11B"
)
markers <- markers
for (i in seq_along(remove)) {
  markers <- rapply(
    markers, function(x) ifelse(x == remove[i], "", x), how = "replace")
}
markers <- lapply(markers, function(x) x[!(x == "")])
remove <- which(
  names(markers) %in% c(
    "Radial Glia_vRG", "Radial Glia_tRG", "Microglia", "oRG specific genes", "vRG specific genes"
  )
)
markers <- markers[-remove]

final_annotations <- c()
final_annotations_long <- c()
cell_order <- c()
```

# Excitatory Neuron

## Standard Pipeline

```{r}
sub_seurat <- seurat[ , which(seurat$cluster_names == "Excitatory Neuron")]

DefaultAssay(sub_seurat) <- "RNA"
sub_seurat[["SCT"]] <- NULL
sub_name <- "Ex"
sub_seurat <- cluster_pipeline(
  sub_seurat, cache_dir, sub_name, protocol, vars_to_regress, parallel_override, FALSE, 15000
)
sub_seurat

for (i in 1:length(metadata_to_plot)) {
  print(names(metadata_to_plot)[i])
  if ("no_legend" %in% metadata_to_plot[[i]] && "no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i]) + NoLegend())
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat") + NoLegend()
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i]))
  } else {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat"))
  }
}

for (j in seq_along(markers)) {
  if (any(markers[[j]] %in% rownames(sub_seurat))) {
    print(
      FeaturePlot(
        sub_seurat, markers[[j]],
        order = TRUE, cols = c("lightgrey", "red")
      ) +
        plot_annotation(
          names(markers)[j],
          theme = theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"))
        )
    )
  }
}
```

# Radial Glia

## Standard Pipeline

```{r}
sub_seurat <- seurat[ , which(seurat$cluster_names == "Radial Glia")]

DefaultAssay(sub_seurat) <- "RNA"
sub_seurat[["SCT"]] <- NULL
sub_name <- "RG"
sub_seurat <- cluster_pipeline(
  sub_seurat, cache_dir, sub_name, protocol, vars_to_regress, parallel_override, FALSE, 15000
)
sub_seurat

for (i in 1:length(metadata_to_plot)) {
  print(names(metadata_to_plot)[i])
  if ("no_legend" %in% metadata_to_plot[[i]] && "no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i]) + NoLegend())
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat") + NoLegend()
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i]))
  } else {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat"))
  }
}

for (j in seq_along(markers)) {
  if (any(markers[[j]] %in% rownames(sub_seurat))) {
    print(
      FeaturePlot(
        sub_seurat, markers[[j]],
        order = TRUE, cols = c("lightgrey", "red")
      ) +
        plot_annotation(
          names(markers)[j],
          theme = theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"))
        )
    )
  }
}
```

# Inhibitory Neuron

## Standard Pipeline

```{r}
sub_seurat <- seurat[ , which(seurat$cluster_names == "Inhibitory Neuron")]

DefaultAssay(sub_seurat) <- "RNA"
sub_seurat[["SCT"]] <- NULL
sub_name <- "In"
sub_seurat <- cluster_pipeline(
  sub_seurat, cache_dir, sub_name, protocol, vars_to_regress, parallel_override, FALSE, 15000
)
sub_seurat

for (i in 1:length(metadata_to_plot)) {
  print(names(metadata_to_plot)[i])
  if ("no_legend" %in% metadata_to_plot[[i]] && "no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i]) + NoLegend())
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat") + NoLegend()
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i]))
  } else {
    print(red_dim_plot(sub_seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat"))
  }
}

for (j in seq_along(markers)) {
  if (any(markers[[j]] %in% rownames(sub_seurat))) {
    print(
      FeaturePlot(
        sub_seurat, markers[[j]],
        order = TRUE, cols = c("lightgrey", "red")
      ) +
        plot_annotation(
          names(markers)[j],
          theme = theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"))
        )
    )
  }
}
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
